---
title: "R : Project 7 : Census Income"
output:
  html_document:
    df_print: paged
header-includes: \usepackage{fvextra} \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      tidy = 'formatR',
                      #tidy.opts = list(blank = FALSE, width.cutoff = 90),
                      prompt = F,
                      size = 'large')
```

### Loading required packages

```{r}
library('dplyr')
library('caTools')
library('ggplot2')
library('caret')
library('stringr')
library('ROCR')
library('randomForest')
library('rpart')
library('rpart.plot')
```

### Loading the data-set

```{r}
census <- read.csv("/home/carb0n/Desktop/r/d/p7/census-income.csv") 
```


### Preparing the data

#### Looking at the structure of the data-set

```{r}
str(census)
```

#### Replacing " ?" with NA for character type columns

```{r}
census[census == " ?"] <- NA
```

#### Checking for null values

```{r}
sum(is.na(census))
```

#### Removing all NA values

```{r}
census <- na.omit(census)
```

#### Removing white spaces for character type columns

```{r}
census <- census %>% mutate_if(is.character, str_trim)
```

#### Converting all character type columns to factor type

```{r}
census <- census %>% mutate_if(sapply(census, is.character), as.factor)
```

### Data Manipulation

#### Extracting education column and storing in census_ed

```{r}
census_ed <- census %>% select(education) 
head(census_ed)
```

#### Extracting all columns from "age" to "relationship" and storing in census_seq

```{r}
census_seq <- census %>% select(age:relationship)
head(census_seq)
```

#### Extracting columns 5, 8, 11 and storing in census_col

```{r}
census_col <- census %>% select(5,8,11)
head(census_col)
```

#### Extracting all the "Male" employees who work in state-gov and storing in male_gov

```{r}
male_gov <- census %>% select(sex,workclass) %>% filter(sex=="Male" & workclass=="State-gov")
head(male_gov)
```

#### Extracting all the 39 year old who either have bachelor's degree or are native of U.S. and storing in census_us

```{r}
census_us <- census %>% select(age,education,native.country) %>% filter(age==39,education=="Bachelors"|native.country=="United-States")

head(census_us)
```

#### Extracting 200 random rows from the “census” data frame and storing it in “census_200”.

```{r}
census_200 <- census %>% sample_n(200) 
head(census_200)
```

#### Count for different levels of "workclass" column

```{r}
census %>% count(workclass)
```

#### Calculating the mean of "capital gain" column and grouping it accoring to "workclass" column

```{r}
census %>% group_by(workclass) %>% summarise(mean(capital.gain))
```

### Data Visualization

#### (A) BAR PLOT

```{r}
# Bar-plot for the “relationship” column and filled the bars according to the “race” column.
# Fill the bars according to "sex"
# Set the Position of the bars to "dodge"
# Set title of the plot to be "Distribution of Relationships by Sex"

ggplot(data = census, aes(x = relationship, fill = sex))+
  geom_bar(position = "dodge")+
  labs(x = "Categories of Relationships", y = "Count of Categories",
       title = "Distribution of Relationships by Sex")
```


#### (B) HISTOGRAM

```{r}
# Histogram for "age" column with number of bins = 50, fill with yearly income (X)
# Set title to be Distribution of Age
# Set legend title to be "Yearly Income"
# Set theme of the plot to be black and white

ggplot(data = census, aes(x = age, fill = X))+
  geom_histogram(bins = 50)+
  labs(title = "Distribution of Age", fill = "Yearly Income")+
  theme_bw()
```

#### (C) SCATTER PLOT

```{r}
# Scatter plot between capital gain and hours per week
# Set Transparency of the points to 40% and size as 2
# Set color of the points according to X (Yearly Income) column
# Set labels

ggplot(data = census, aes(x = capital.gain, y = hours.per.week, col = X))+
  geom_point(alpha = 0.6, size = 2)+
  labs(x = "Capital Gain", y = "Hours Per Week", col = "Yearly Income")
```

#### (D) BOX PLOT

```{r}
# Box plot between eucation and age column
# Fill according to sex column
# Set title to "Box plot of age by education and sex)

ggplot(data = census, aes(x = education, y = age, fill = sex))+
  geom_boxplot()+
  labs(title = "Box-Plot of Age by Education and Sex")
```

### (1) Linear Regression

### Splitting the data-set

```{r}
set.seed(1)
sample_tag <- sample.split(census$hours.per.week, SplitRatio = 0.7)
train_data <- subset(census, sample_tag==TRUE)
test_data <- subset(census, sample_tag==FALSE)

nrow(train_data)
nrow(test_data)
```

#### Building Linear Regression Model

```{r}
lr_model <- lm(hours.per.week ~ education.num, data = train_data)
```

#### Checking Summary of the Linear Regression Model

```{r}
summary(lr_model)
```

#### Predicting on test data-set using linear regression model

```{r}
lr_pred_val <- predict(lr_model, newdata = test_data) 
```

#### Checking predicted values for linear regression model

```{r}
head(lr_pred_val)
```

#### Binding predicted values with actual values to find error

```{r}
lr_final_data <- cbind(Actual = test_data$hours.per.week, Predicted = lr_pred_val)
lr_final_data <- as.data.frame(lr_final_data)
head(lr_final_data)
```

#### Finding RMSE (root-mean-squared-error)

```{r}
lr_final_data <- lr_final_data %>% mutate(error = lr_final_data$Actual - lr_final_data$Predicted)
sqrt(mean(lr_final_data$error)^2)

# RMSE = 0.03315308
# Accuracy = 99.97%
```

### (2) Simple Logistic Regression Model

#### Splitting the data-set

```{r}
set.seed(1)
sample_tag <- sample.split(census$X, SplitRatio = 0.65)
train_data <- subset(census, sample_tag==TRUE)
test_data <- subset(census, sample_tag==FALSE)

nrow(train_data)
nrow(test_data)
```

#### Building Logistic Regrssion Model

```{r}
log_model_1 <- glm(X ~ occupation, data = train_data, family = "binomial")
```

#### Checking summary of the model

```{r}
summary(log_model_1)
```

#### Predicting on test-data

```{r}
log_pred_val_1 <- predict(log_model_1, newdata = test_data, type = "response") 
```

#### Checking predicted values 

```{r}
head(log_pred_val_1)
```

#### Checking range of predicted values

```{r}
range(log_pred_val_1)
```

#### Plotting Accuracy v/s cut-off and finding an ideal value for cut-off

```{r}
predict_log_acc_1 <- prediction(log_pred_val_1, test_data$X)
```

#### Using performance() to check accuracy 

```{r}
log_acc_1 <- performance(predict_log_acc_1, "acc")
```

#### Plotting "log_acc_1" to find value where accuracy is constant

```{r}
plot(log_acc_1)
```

##### Ideal cut-off (threshold) value based on plot => 0.47

#### Checking range of predicted values

```{r}
range(log_pred_val_1)
```

#### Checking distribution of dependent variable

```{r}
table(census$X)
```

#### Finding Accuracy using threshold

```{r}
log_pred_1 <- ifelse(log_pred_val_1 > 0.47, ">50K", "<=50K")
cm_log_1 <- table(log_pred_1, test_data$X)
cm_log_1

sum(diag(cm_log_1))/sum(cm_log_1)
# Accuracy = 74%
```

#### Building confusion matrix  and finding accuracy

```{r}
confusionMatrix(factor(log_pred_1),test_data$X)
# Accuracy = 74%
```

#### Plotting ROC Curve

```{r}
log_roc_1 <- performance(predict_log_acc_1, "tpr", "fpr")
plot(log_roc_1)
```

#### Finding AUC (Area Under the Curve)

```{r}
log_auc_1 <- performance(predict_log_acc_1, "auc")
log_auc_val_1 <- log_auc_1@y.values[[1]]
log_auc_val_1
# AUC = 0.7256113
```

### (3) Multiple Logistic Regression Model

#### Splitting the data-set

```{r}
set.seed(1)
sample_tag <- sample.split(census$X, SplitRatio = 0.80)
train_data <- subset(census, sample_tag==TRUE)
test_data <- subset(census, sample_tag==FALSE)

nrow(train_data)
nrow(test_data)
```

#### Building Logistic Regression Model

```{r}
log_model_2 <- glm(X ~ age+workclass+education, data = train_data, family = "binomial")
```

#### Checking summary of the Logistic Regression Model

```{r}
summary(log_model_2)
```

#### Predicting on test-data 

```{r}
log_pred_val_2 <- predict(log_model_2, newdata = test_data, type = "response") 
```

#### Checking predicted values 

```{r}
head(log_pred_val_2)
```

#### Checking range of predicted values

```{r}
range(log_pred_val_2)
```

#### Plotting Accuracy v/s cut-off and finding an ideal value for cut-off

```{r}
predict_log_acc_2 <- prediction(log_pred_val_2, test_data$X)
```

#### Using performance() to check accuracy

```{r}
log_acc_2 <- performance(predict_log_acc_2, "acc")
```

#### Plotting "log_acc_1" to find value where accuracy is constant

```{r}
plot(log_acc_2)
```

##### Ideal cut-off value based on plot => 0.45

#### Checking distribution of dependent variable 

```{r}
table(census$X)
```

#### Setting threshold for prediction

```{r}
log_pred_2 <- ifelse(log_pred_val_2 > 0.45, ">50K", "<=50K")
```

#### Finding accuracy

```{r}
cm_log_2 <- table(log_pred_2, test_data$X)
cm_log_2

sum(diag(cm_log_2))/sum(cm_log_2)
```

#### Building confusion-matrix and finding accuracy

```{r}
confusionMatrix(factor(log_pred_2),test_data$X)
```

#### Plotting the ROC Curve

```{r}
log_roc_2 <- performance(predict_log_acc_2, "tpr", "fpr")
plot(log_roc_2)
```

#### Finding AUC (Area Under the Curve)

```{r}
log_auc_2 <- performance(predict_log_acc_2, "auc")
log_auc_val_2 <- log_auc_2@y.values[[1]]
log_auc_val_2
# AUC = 0.7741555
```

### (4) Decision Tree

#### Splitting the data-set

```{r}
set.seed(1)
sample_tag <- sample.split(census$X, SplitRatio = 0.70)
train_data <- subset(census, sample_tag==TRUE)
test_data <- subset(census, sample_tag==FALSE)

nrow(train_data)
nrow(test_data)
```

#### Building decision tree model

```{r}
census_dt <- rpart(formula = X ~ .,
                   data = train_data,
                   method = "class")
```

#### Plotting the decision tree

```{r}
rpart.plot(x = census_dt, type = 5, extra = 0, tweak = 1.2)
```

#### Predicting on test-data

```{r}
class_pred_dt <- predict(object = census_dt,
                      newdata = test_data,
                      type = "class")
```

#### Finding Accuracy

```{r}
table_dt <- table(Prediction = class_pred_dt, Actual = test_data$X) 
sum(diag(table_dt))/sum(table_dt)
# Accuracy = 0.8389
```

#### Building Confusion Matrix and Finding Accuracy

```{r}
confusionMatrix(class_pred_dt, test_data$X)
# Accuracy = 84%
```

### (5) Random Forest

#### Splitting the data-set

```{r}
set.seed(1)
sample_tag <- sample.split(census$X, SplitRatio = 0.80)
train_data <- subset(census, sample_tag==TRUE)
test_data <- subset(census, sample_tag==FALSE)

nrow(train_data)
nrow(test_data)
```

#### Building random-forest model

```{r}
census_rf <- randomForest(formula = X ~ .,
                          data = train_data,
                          ntree = 300)
```

#### Printing the random-forest model details

```{r}
print(census_rf)
```

#### Plotting the random-forest model

```{r}
plot(census_rf)
```

#### Predicting on test-data

```{r}
class_pred_rf <- predict(object = census_rf,
                         newdata = test_data,
                         type = "class")
```

#### Finding Accuracy

```{r}
table_rf <- table(Prediction = class_pred_rf, Actual = test_data$X)
table_rf

sum(diag(table_rf))/sum(table_rf)
# Accuracy = 85.7%
```

#### Building confusion-matrix and finding accuracy

```{r}
confusionMatrix(class_pred_rf, test_data$X)
```