---
title: "R : Project 5 : Loan Approval Prediction"
output:
  html_document:
    df_print: paged
header-includes: \usepackage{fvextra} \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      tidy = 'formatR',
                      #tidy.opts = list(blank = FALSE, width.cutoff = 90),
                      prompt = F,
                      size = 'large')
```

### Loading required packages

```{r}
library('dplyr')
library('e1071')
library('caret')
library('caTools')
```

### Loading the data-set

```{r}
loan <- read.csv("/home/carb0n/Desktop/r/d/p5/customer_loan.csv",stringsAsFactors = TRUE)
```

### Preparing the data

#### Looking at the structure of data-set

```{r}
str(loan)
```

#### Checking for null values in data-set

```{r}
sum(is.na(loan))
```

#### Adding new column "dti"

```{r}
loan <- loan %>% mutate(dti = debts/income)
```

#### Adding new column "loan_decision_status"

```{r}
loan <- loan %>% mutate(loan_decision_status = if_else(loan_decision_type == "Denied", 0, 1))
```

#### Converting column "loan_decision_status" to factor type

```{r}
loan$loan_decision_status <- as.factor(loan$loan_decision_status)
```

#### Creating new data-frame using selected columns

```{r}
customer_loan_refined <- loan %>% select(3,4,6,7,8,11,13,14)
```

#### Converting appropriate variables to numeric type

```{r}
customer_loan_refined$gender <- as.numeric(customer_loan_refined$gender)
customer_loan_refined$marital_status <- as.numeric(customer_loan_refined$marital_status)
customer_loan_refined$occupation <- as.numeric(customer_loan_refined$occupation)
customer_loan_refined$loan_type <- as.numeric(customer_loan_refined$loan_type)
```

### Splitting the data-set

#### Creating train and test sets

```{r}
set.seed(1)
split_tag <- sample.split(customer_loan_refined$loan_decision_status, SplitRatio = 0.7)
train_data <- subset(customer_loan_refined,split_tag==TRUE)
test_data <- subset(customer_loan_refined,split_tag==FALSE)
```

### Feature Scaling

```{r}
# Scaling features from train-data
train_scale <- scale(train_data[,1:7])

# Scaling features from test-data
test_scale <- scale(test_data[,1:7])
```

## PCA (Principal Component Analysis)

### (1) PCA (On Train-Data)

#### Performing Principal Component Analysis (PCA) on train-data

```{r}
train_pca <- prcomp(train_data[,c(1:7)], center = TRUE, scale. = TRUE)
```

#### Checking summary of PCA performed on train-data

```{r}
summary(train_pca)
```

##### Based on summary analysis, PC-6 and PC-7 are of higher importance in train-data.

### (2) PCA (On Test-Data)

#### Performing PCA on test-data

```{r}
test_pca <- prcomp(test_data[,c(1:7)], center = TRUE, scale. = TRUE)
```

#### Checking summary of PCA performed on test-data

```{r}
summary(test_pca)
```

##### Based on summary analysis, PC-6 and PC-7 are of higher importance in test-data.

### Naive-Bayes - Model Building, Prediction, Confusion-Matrix And Evaluation

#### Building Naive Bayes Model

```{r}
classifier_loan <- naiveBayes(loan_decision_status ~ ., data = train_data)
```

#### Predicting on test data-set

```{r}
classifier_loan_pred <- predict(classifier_loan, newdata = test_data)
```

#### Building Confusion Matrix

```{r}
cm <- table(test_data$loan_decision_status, classifier_loan_pred)
```

#### Model Evaluation

```{r}
confusionMatrix(cm)
```