---
title: "R : Project 2 : Recommendation Engine (Book Recommender)"
output:
  html_document:
    df_print: paged
header-includes: \usepackage{fvextra} \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      tidy = 'formatR',
                      #tidy.opts = list(blank = FALSE, width.cutoff = 90),
                      prompt = F,
                      size = 'large')
```

### Loading required packages 

```{r}
library(recommenderlab)
library(data.table)
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
library(DT)
library(knitr)
library(grid)
library(gridExtra)
library(corrplot)
library(qgraph)
library(methods)
library(Matrix)
```


### Loading the data-sets

```{r}
books <- read.csv("/home/carb0n/Desktop/r/d/p2/books.csv")
ratings <- read.csv("/home/carb0n/Desktop/r/d/p2/ratings.csv")
book_tags <- read.csv("/home/carb0n/Desktop/r/d/p2/book_tags.csv")
tags <- read.csv("/home/carb0n/Desktop/r/d/p2/tags.csv")
```

### Viewing the data-sets

```{r}
#View(ratings)
#View(books)
#View(book_tags)
#View(tags)
```

### Preparing the data
  
#### Finding and removing duplicate ratings

```{r}
# Finding count of different ratings
ratings %>% group_by(user_id, book_id) %>% mutate(N=n()) -> ratings
table(ratings$N)

# Filtering duplicate ratings
ratings %>% filter(N>1) -> duplicate_ratings  

# Removing duplicate ratings
ratings %>% filter(N==1) -> ratings
```

#### Removing users who rated fewer than 3 books

```{r}
# Finding count of times a user has provided ratings
ratings %>% group_by(user_id) %>% mutate(Ratings_Given = n()) -> ratings

# Keeping only those users who rated more than 2 times
ratings %>% filter(Ratings_Given>2) -> ratings
```

### Sampling the data

#### Selecting a sample from the entire data-set

```{r}
set.seed(1)
user_fraction <- 0.02
users <- unique(ratings$user_id)
sample_users <- sample(users, round(user_fraction * length(users)))  
ratings %>% filter(user_id %in% sample_users) -> ratings
```

### Exploring the data

#### Plotting distribution of ratings 

```{r}
ratings %>% 
  ggplot(aes(x = rating, fill = factor(rating))) +
  geom_bar(color = "grey20") + scale_fill_brewer(palette = "YlGnBu") + guides(fill = FALSE)
```

#### Plotting number of ratings per book

```{r}
ratings %>% 
  group_by(book_id) %>% 
  summarize(number_of_ratings_per_book = n()) %>% 
  ggplot(aes(number_of_ratings_per_book)) + 
  geom_bar(fill = "orange", color = "grey20", width = 1) + coord_cartesian(c(0,40))
```

#### Finding the count of different genres

```{r}
genres <- str_to_lower(c("Art", "Biography", "Business", "Chick Lit", "Children's", "Christian", "Classics", "Comics", "Cookbooks", "Crime", "Fantasy", "Gay and Lesbian", "Graphic Novels", "Historical Fiction", "History", "Horror", "Humor and Comedy", "Manga", "Memoir", "Music", "Mystery", "Paranormal", "Philosophy", "Poetry", "Psychology", "Religion", "Romance", "Science", "Science Fiction", "Self Help", "Suspense", "Spirituality", "Sports", "Thriller", "Travel", "Young Adult"))

available_genres <- genres[str_to_lower(genres) %in% tags$tag_name]
available_genres

available_tags <- tags$tag_id[match(available_genres, tags$tag_name)]
head(available_tags)
```

#### Plotting the percentage of each genre

```{r}
tmp <- book_tags %>% 
  filter(tag_id %in% available_tags) %>% 
  group_by(tag_id) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  mutate(sumN = sum(n), percentage = n / sumN) %>%
  arrange(-percentage) %>%
  left_join(tags, by = "tag_id")  
```
```{r}
tmp %>% 
  ggplot(aes(reorder(tag_name, percentage), percentage, fill = percentage)) + geom_bar(stat = "identity") + coord_flip() + scale_fill_distiller(palette = 'YlOrRd') + labs(y = 'Percentage', x = 'Genre')

```

#### Finding the top 10 books with highest ratings

```{r}
books %>% 
  arrange(-average_rating) %>% 
  top_n(10,wt = average_rating) %>% 
  select(title, ratings_count, average_rating) 
```

#### Finding the 10 most popular books

```{r}
books  %>% 
  arrange(-ratings_count) %>% 
  top_n(10,wt = ratings_count) %>% 
  select(title, ratings_count, average_rating) 
```

### Preparing the data

#### Reshaping the data to build collaborative filtering model

```{r}
dimension_names <- list(user_id = sort(unique(ratings$user_id)), book_id = sort(unique(ratings$book_id)))
ratingmat <- spread(select(ratings, book_id, user_id, rating), book_id, rating) %>% select(-user_id)
```

#### Preparing a rating matrix

```{r}
# Converting ratingmat to matrix
ratingmat <- as.matrix(ratingmat)

# Removing "user_id" column
ratingmat[,-1] -> ratingmat

# Providing dimension names as mentioned in dimension_names
dimnames(ratingmat) <- dimension_names

# Checking dimensions of rating-matrix
dim(ratingmat)

# Viewing first five records from rating-matrix
ratingmat[1:5, 1:5]
```

#### Preparing a realRatingMatrix

```{r}
# Creating another rating-matrix from original rating-matrix
ratingmat0 <- ratingmat  

# Replacing all NA values with '0'
ratingmat0[is.na(ratingmat0)] <- 0

# Converting rating-matrix to sparseMatrix
sparse_ratings <- as(ratingmat0, "sparseMatrix")

# Converting sparse-matrix to realRatingMatrix
real_ratings <- new("realRatingMatrix", data = sparse_ratings)
```

### Splitting the data into train & test sets

```{r}
sample(x=c(T,F),size=nrow(real_ratings),replace = T, prob = c(0.8,0.2)) ->split_book
real_ratings[split_book,]->recc_train
real_ratings[!split_book,]->recc_test
```

### Model Building and Prediction

#### Building User-Based Collaborative Filtering (UBCF) Model

```{r}
Recommender(data = recc_train,method="UBCF")->recc_model_ubcf
```

#### Setting number of desired recommendations

```{r}
n_recommended_ubcf<-6
```

#### Predicting for User-1 from test-data

```{r}
predict(object=recc_model_ubcf,newdata=recc_test[1],n=n_recommended_ubcf)->recc_predicted_ubcf
```

#### Recommending books for User-1

```{r}
recc_predicted_ubcf@items[[1]]->user1_book_numbers
recc_predicted_ubcf@itemLabels[user1_book_numbers]
```

#### Viewing recommended books for User-1

```{r}
books %>% filter(id==6343) %>% select(original_title,authors)
books %>% filter(id==7482) %>% select(original_title,authors)
books %>% filter(id==2750) %>% select(original_title,authors)
```

#### Predicting for User-5 in test-data

```{r}
predict(object=recc_model_ubcf,newdata=recc_test[5],n=n_recommended_ubcf)->recc_predicted_ubcf
```

#### Recommending books for User-5

```{r}
recc_predicted_ubcf@items[[1]]->user5_book_numbers
recc_predicted_ubcf@itemLabels[user5_book_numbers]
```

#### Viewing recommended books for User-5

```{r}
books %>% filter(id==4624) %>% select(original_title,authors)
books %>% filter(id==6867) %>% select(original_title,authors)
books %>% filter(id==7326) %>% select(original_title,authors)
```