---
title: "R : Project 6 : Netflix Recommender System Using Recommenderlab"
output:
  html_document:
    df_print: paged
header-includes: \usepackage{fvextra} \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      tidy = 'formatR',
                      #tidy.opts = list(blank = FALSE, width.cutoff = 90),
                      prompt = F,
                      size = 'large')
```

### Loading required packages

```{r}
library('dplyr')
library('reshape2')
library('ggplot2')
library('recommenderlab') 
```

### Loading the netflix data-set

```{r}
data1 <- read.csv("/home/carb0n/Desktop/r/d/p6/combined_data_1.txt", sep = ",", header = FALSE, stringsAsFactors = FALSE)
```

### Preparing the data

#### Viewing first few records from beginning
```{r}
head(data1)
```

#### Checking structure of the data-set

```{r}
str(data1)
```

#### Removing column "" as it is not needed

```{r}
data1 <- data1[,-3]
```

#### Renaming columns

```{r}
colnames(data1) <- c("Cust_Id", "Rating")
head(data1)
```

#### Checking for NA values

```{r}
sum(is.na(data1))
```

### Exploring the data

#### Total count of different ratings in the data-set

```{r}
data <- data1 %>% group_by(Rating) %>% summarise(count = n())

#### Removing the row showing count of NA values
data <- data[-6,]

#### Viewing count of different ratings
head(data)
```

#### Movie-Count by counting NaN values

```{r}
movie_count <- sum(is.na(data1))
```

#### Unique customer count 

```{r}
cust_count <- data1 %>% summarise(n_distinct(Cust_Id)) - movie_count

# Converting obtained count to integer format
cust_count <- as.integer(cust_count)

# Viewing unique customer count
cust_count
```

#### Total count of ratings given

```{r}
rating_count <- (nrow(data1)) - movie_count

# Viewing total count of ratings given
rating_count
```

### Plotting the data

#### To disable the scientific notation on axis-scale and instead show actual value

```{r}
options(scipen = 999)
```

#### Plotting total count of rated movies, ratings and customers

```{r}
ggplot(data, aes(factor(Rating), y = count, fill = factor(Rating)))+
  geom_bar(stat = "identity")+
  geom_text(aes(label = sprintf("%d%%", round(count/sum(count) * 100))), vjust = -0.5)+
  labs(x = "Rating", y = "Rating Count", fill = "Rating")+
  labs(title = sprintf("Total Pool: %d Movies, %d Customers, %d Ratings Given\n",
                       movie_count, cust_count, rating_count))+
  theme_bw()
```


## Preprocessing the data

### Loading the modified netflix data-set

```{r}
data <- read.csv("/home/carb0n/Desktop/r/d/p6/netflix_data.csv", stringsAsFactors = FALSE)
head(data)
```

#### Removing column "X" as it is not needed

```{r}
data <- data[,-1]
```

### Loading the movie-titles data-set

```{r}
movies <- read.csv("/home/carb0n/Desktop/r/d/p6/movie_titles.csv", encoding = "ISO-8859-1", header = FALSE, stringsAsFactors = FALSE)
```

#### Renaming the columns

```{r}
colnames(movies) <- c("Movie_Id", "Year", "Name")
```

#### Checking the structure of the data-set
```{r}
str(movies)
```

### Preparing the data

#### Grouping movies by their ratings count and mean rating

```{r}
df_movie_summary <- data %>% group_by(Movie_Id) %>% summarise(count = n(), mean(Rating))
```

#### Setting benchmark for including movies

```{r}
movie_benchmark <- round(quantile(df_movie_summary$count,0.7,names = FALSE),0)
```

#### Creating list of movies that are above benchmark

```{r}
drop_movie_list <- df_movie_summary[df_movie_summary$count < movie_benchmark,1]
```

#### Converting obtained list of movies to data-frame

```{r}
drop_movie_list <- as.data.frame(drop_movie_list)
```

#### Viewing list 

```{r}
head(drop_movie_list)
```

#### Grouping movies by count of ratings given and customer-id

```{r}
df_movie_summary1 <- data %>% group_by(Cust_Id) %>% summarise(count = n(),mean(Cust_Id))
```

#### Setting benchmark for including customer-id

```{r}
cust_benchmark <- round(quantile(df_movie_summary1$count,0.7,names = FALSE),0)
```

#### Creating list of customer-ids that are above benchmark

```{r}
drop_cust_list <- df_movie_summary1[df_movie_summary1$count < cust_benchmark,1]
```

#### Converting obtained list of movies to data-frame

```{r}
drop_cust_list <- as.data.frame(drop_cust_list)
```

#### Viewing list 

```{r}
head(drop_cust_list)
```

#### Removing Movies that don't satisfy the benchmark

```{r}
data <- anti_join(data, drop_movie_list, by = "Movie_Id")
```

#### Removing Customer IDs that don't satisfy the benchmark

```{r}
data <- anti_join(data, drop_cust_list, by = "Cust_Id")
```

#### Viewing updated data

```{r}
head(data)
```

### Preparing the ratingmatrix for recommenderlab

#### Creating Rating Matrix where Rows = UserId and Columns = MovieId

```{r}
ratingmat <- dcast(data, Cust_Id~Movie_Id, value.var = "Rating", na.rm = FALSE)
```

#### Removing column "Cust_Id" 

```{r}
ratingmat <- as.matrix(ratingmat[,-1]) 
```

#### Converting rating matrix into recommenderlab sparse matrix

```{r}
ratingmat <- as(ratingmat, "realRatingMatrix")
```

#### Normalizing the data

```{r}
ratingmat_norm <- normalize(ratingmat)
```

### UBCF(User-Based-Collaborative-Filtering) Model Creation and Prediction

#### Creating recommender model using "User-Based-Collaborative-Filtering" (UBCF) method

```{r}
recommender_model <- Recommender(ratingmat_norm, method = "UBCF", param = list(method = "Cosine",nn = 30))
```

#### Predicting for User-1 in test-data

```{r}
recom <- predict(recommender_model, ratingmat[1], n = 10)
```

#### Putting obtained recommendations for User-1 in test-data in list format

```{r}
recom_list <- as(recom, "list")
```

### Recommendations

#### Creating an empty matrix to save recommended movie predictions

```{r}
recom_result <- matrix(0,10)
```

#### Obtaining top-3 recommendations for User-1 in test-data 

```{r}
for(i in c(1:10)){
  recom_result[i] <- movies[as.integer(recom_list[[1]][i]),3]
}
```

#### Viewing top-3 recommendations for User-1 in test-data 

```{r}
recom_result
```

### Evaluation

#### Creating evaluationscheme with 3-fold cross-validation to evaluate the rating matrix

```{r}
evaluation_scheme <- evaluationScheme(ratingmat, method = "cross-validation", train = 0.8, k = 3, given = -1, goodRating = 5)
```

#### Performing evaluation using SVD (Single Value Decomposition) method

```{r}
evaluation_result <- evaluate(evaluation_scheme, method = "SVD", n = c(1,3,5,10))
```

#### Building confusion-matrix of evaluation results

```{r}
eval_result <- getConfusionMatrix(evaluation_result)[[1]]
```

#### Viewing confusion-matrix of evaluation results

```{r}
eval_result
```