---
title: "R : Project 1 : Market-Basket-Analysis"
output:
  html_document:
    df_print: paged
header-includes: \usepackage{fvextra} \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      tidy = 'formatR',
                      #tidy.opts = list(blank = FALSE, width.cutoff = 90),
                      prompt = F,
                      size = 'large')
```

### Loading required packages

```{r}
library(dplyr)
library(arules)
library(arulesViz)
library(plotly)
library(RColorBrewer)
```

### Exploring the data

```{r}
# Loading the data-set
retail <- read.csv("/home/carb0n/Desktop/r/d/p1/market_basket.csv")

# Loading the data-set
tr <- read.transactions("/home/carb0n/Desktop/r/d/p1/market_basket.csv", format="basket", sep=",")

# Checking summary of the data-set
summary(tr)
```

### Summary Related Answers:

##### Total number of transactions => 18841 
##### Total number of items in the inventory => 26230
##### Total number of items purchased =>

```{r}
18441*26230*0.0007227055  
```

##### Plot for identifying 10 most frequently bought items

```{r}
itemFrequencyPlot(tr, topN=10, type="absolute", col=brewer.pal(8,'Pastel2'),
                  main="Absolute Item Frequency Plot")
```

### (A) 1st Set of Association Rules

#### Creation of 1st Set of Association Rules

```{r}
# Using min support as 0.005 and confidence as 0.8.
association.rules_1 <- apriori(tr, parameter = list(supp=0.005, conf=0.8,maxlen=10))

# Checking summary of association rule-1
summary(association.rules_1)
```

#### Inspection of 1st Set Of Association Rules

```{r}
# Inspection of top 5 rules sorted w.r.t. confidence  
inspect(head(sort(association.rules_1,by='confidence'),5))

# Inspection of bottom 5 rules sorted w.r.t. confidence
inspect(tail(sort(association.rules_1,by='confidence'),5))

# Inspection of top 5 rules sorted w.r.t. lift
inspect(head(sort(association.rules_1,by='lift'),5))
```

#### Plotting 1st Set of Association Rules

```{r}
# Plotting all the rules
plot(association.rules_1, method='scatter', engine='html')
```

#### Rules with confidence greater than 0.4 or 40% (from 1st set)

```{r}
# Filter rules with confidence greater than 0.4 or 40%
subRules_ar1 <- association.rules_1[quality(association.rules_1)$confidence>0.4]

# Interactive scatter plot for subRules_ar1 using plotly engine
plot(subRules_ar1, method="scatter", engine="plotly")

# Two-key plot for subRules_ar1
plot(subRules_ar1, method="two-key plot")
```

#### Top-10 rules with highest "confidence" (from 1st set)

```{r}
# Filter top10 rules with highest confidence
subRules_ar1 <- head(subRules_ar1, n=10, by='confidence')

# Interactive plot for subRules_ar1 using html engine
plot(subRules_ar1, method="graph", engine="htmlwidget")
```

#### Top-10 rules with highest "lift" (from 1st set)

```{r}
# Filter top10 rules with highest lift
subRules_ar1 <- head(subRules_ar1, n=10, by='lift')

# Plot for subRules_ar1
plot(subRules_ar1, method="paracoord", main='Top10 rules with highest lift')
```

### (B) 2nd Set of Association Rules

#### Creation of 2nd Set of Association Rules 

```{r}
# Using min support as 0.009 and confidence as 0.3.
association.rules_2 <- apriori(tr, parameter = list(supp=0.009, conf=0.3,maxlen=10))

# Checking summary of association rules-2
summary(association.rules_2)
```

#### Inspection of 2nd Set of Association Rules

```{r}
# Inspection of top 5 rules sorted w.r.t. confidence  
inspect(head(sort(association.rules_2,by='confidence'),5))

# Inspection of bottom 5 rules sorted w.r.t. confidence
inspect(tail(sort(association.rules_2,by='confidence'),5))

# Inspection of top 5 rules sorted w.r.t. lift
inspect(head(sort(association.rules_2,by='lift'),5))
```

#### Plotting 2nd Set of Association Rules

```{r}
# Plotting all the rules
plot(association.rules_2, method='scatter', engine='html')
```

#### Rules with confidence greater than 0.4 or 40% (from 2nd set)

```{r}
# Filter rules with confidence greater than 0.4 or 40%
subRules_ar2 <- association.rules_2[quality(association.rules_2)$confidence>0.4]

# Interactive scatter plot for subRules_ar1 using plotly engine
plot(subRules_ar2, method="scatter", engine="plotly")

# Two-key plot for subRules_ar1
plot(subRules_ar2, method="two-key plot")
```

#### Top-10 rules with highest "confidence" (from 2nd set)

```{r}
# Filter top10 rules with highest confidence
subRules_ar2 <- head(subRules_ar2, n=10, by='confidence')

# Interactive plot for subRules_ar1 using html engine
plot(subRules_ar2, method="graph", engine="htmlwidget")
```

#### Top-10 rules with highest "lift" (from 2nd set)

```{r}
# Filter top10 rules with highest lift
subRules_ar2 <- head(subRules_ar2, n=10, by='lift')

# Plot for subRules_ar1
plot(subRules_ar2, method="paracoord", main='Top10 rules with highest lift')
```

### (C) 3rd Set of Association Rules

#### Creation of 3rd Set of Association Rules

```{r}
# Using min support as 0.002 and confidence as 0.5.
association.rules_3 <- apriori(tr, parameter = list(supp=0.02, conf=0.5,maxlen=10))

# Checking summary of association rules-3
summary(association.rules_3)
```

#### Inspection of 3rd Set of Association Rules
```{r}
# Inspection of top 5 rules sorted w.r.t. confidence  
inspect(head(sort(association.rules_3,by='confidence'),5))

# Inspection of bottom 5 rules sorted w.r.t. confidence
inspect(tail(sort(association.rules_3,by='confidence'),5))

# Inspection of top 5 rules sorted w.r.t. lift
inspect(head(sort(association.rules_3,by='lift'),5))
```

#### Plotting 3rd Set of Association Rules

```{r}
# Plotting all the rules
plot(association.rules_3, method='scatter', engine='html')
```

#### Rules with confidence greater than 0.4 or 40% (from 3rd set)

```{r}
# Filter rules with confidence greater than 0.4 or 40%
subRules_ar3 <- association.rules_3[quality(association.rules_3)$confidence>0.4]

# Interactive scatter plot for subRules_ar1 using plotly engine
plot(subRules_ar3, method="scatter", engine="plotly")

# Two-key plot for subRules_ar1
plot(subRules_ar3, method="two-key plot")
```

#### Top-10 rules with highest "confidence" (from 3rd set)

```{r}
# Filter top10 rules with highest confidence
subRules_ar3 <- head(subRules_ar3, n=10, by='confidence')

# Interactive plot for subRules_ar1 using html engine
plot(subRules_ar3, method="graph", engine="htmlwidget")
```

#### Top-10 rules with highest "lift" (from 3rd set)

```{r}
# Filter top10 rules with highest lift
subRules_ar3 <- head(subRules_ar3, n=10, by='lift')

# Plot for subRules_ar1
plot(subRules_ar3, method="paracoord", main='Top10 rules with highest lift')
```