---
title: "R : Project 3 : Pokemon Case Study"
output:
  html_document:
    df_print: paged
header-includes: \usepackage{fvextra} \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      tidy = 'formatR',
                      #tidy.opts = list(blank = FALSE, width.cutoff = 90),
                      prompt = F,
                      size = 'large')
```

### Loading required packages 

```{r}
library('dplyr')
library('caTools')
library('caret')
library('rpart')
```

### Loading the data-set

```{r}
pokemon <- read.csv("/home/carb0n/Desktop/r/d/p3/Pokemon.csv")
```

### Preparing the data

```{r}
# Checking structure of the data-set
str(pokemon)

# Removing unnecessary columns 
pokemon <- pokemon[,-1]

# Renaming columns
pokemon <- pokemon %>% rename("Primary_Type"=Type.1,"Secondary_Type"=Type.2,"Health_Points"=HP, 
                              "Special_Attack"=Sp..Atk,"Special_Defence"=Sp..Def)

# Converting appropriate variables to factor-type
pokemon$Legendary <- as.factor(pokemon$Legendary)
```

### Exploring the data

#### Looking at different primary types

```{r}
table(pokemon$Primary_Type)
```

#### Looking at different secondary types

```{r}
table(pokemon$Secondary_Type)
```

#### Selecting "Grass" pokemon with "Poison" secondary type and maximum speed

```{r}
Grass_Pokemon <- pokemon %>% filter(Primary_Type=="Grass",Secondary_Type=="Poison") %>% top_n(1,Speed)
Grass_Pokemon
```

#### Selecting "Water" pokemon with "Flying" secondary type and maximum speed

```{r}
Water_Pokemon <- pokemon %>% filter(Primary_Type=="Water",Secondary_Type=="Flying") %>% top_n(1,Defense)
Water_Pokemon
```

#### Selecting "Water" pokemon with "Flying" secondary type and maximum speed

```{r}
Fire_Pokemon <- pokemon %>% filter(Primary_Type=="Fire",Secondary_Type=="Psychic") %>% top_n(1,Attack)
Fire_Pokemon
```

#### Creating data-frame with selected pokemon for each type

```{r}
Selected_Pokemon <- rbind(Grass_Pokemon,Water_Pokemon,Fire_Pokemon) 
Selected_Pokemon
```

## (A) Linear Model

### Splitting the data-set into train and test sets

```{r}
set.seed(1)
split_tag <- sample.split(pokemon$Attack,SplitRatio = 0.65) 
train <- subset(pokemon,split_tag==T)
test <- subset(pokemon,split_tag==F)
```

### Model Building And Prediction

#### Building Linear Model

```{r}
model_1 <- lm(Attack~Defense,data=train)
summary(model_1)
```

#### Predicting the model for test data-set

```{r}
predict_1 <- predict(model_1,newdata=test)
head(predict_1)
```

#### Storing predicted values and actual values in a dataframe

```{r}
final_data <- cbind(Actual=test$Attack,Predicted=predict_1)
final_data <- as.data.frame(final_data)
range(final_data)
head(final_data)
```

### Accuracy And Error

#### Calculating error
```{r}
error <- (final_data$Actual-final_data$Predicted)
final_data <- cbind(final_data,error)
head(final_data)
```

#### Finding RMSE (Root-Mean-Squared Error) 

```{r}
sqrt(mean(final_data$error^2))
```

## (B) Classification-Tree Model

### Splitting the data-set

```{r}
set.seed(1)
split_tag <- sample.split(pokemon$Legendary,SplitRatio = 0.65)
subset(pokemon,split_tag==T) -> train
subset(pokemon,split_tag==F) -> test

nrow(train)
nrow(test)
```

### Preparing the data

```{r}
# Converting "Legendary" column to factor type 
test$Legendary <- as.factor(test$Legendary)

# Converting "Name" column to factor type with same levels as train data
test$Name <- factor(test$Name, levels = levels(train$Name))
```

### Model Building And Prediction

#### Building Classification Tree Model

```{r}
model_2 <- rpart(Legendary~.,data=train)
summary(model_2)
```

#### Predicting on test-data

```{r}
predict_2 <- predict(model_2,newdata=test,type="class")
head(predict_2)
```

#### Confusion-Matrix And Accuracy

```{r}
confusionMatrix(test$Legendary,predict_2)
```