---
title: "R : Project 4 : Credit Card Fraud Detection"
output:
  html_document:
    df_print: paged
header-includes: \usepackage{fvextra} \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      tidy = 'formatR',
                      #tidy.opts = list(blank = FALSE, width.cutoff = 90),
                      prompt = F,
                      size = 'large')
```

### Loading required packages

```{r}
library('dplyr')
library('caTools')
library('caret')
library('ggplot2')
library('rpart')
library('rpart.plot')
library('ROSE')
library('smotefamily')
```

### Loading the data-set

```{r}
cc <- read.csv("/home/carb0n/Desktop/r/d/p4/creditcard.csv")
```

### Preparing the data

#### Looking at structure of data-set

```{r}
str(cc)
```

#### Converting Dependent Variable ("Class") to factor

```{r}
cc$Class <- as.factor(cc$Class)
```

#### Looking at summary of data-set

```{r}
summary(cc)
```

#### Checking for missing values

```{r}
sum(is.na(cc))
```

### Exploring and Analyzing the data (EDA)

#### Distrubution of legitimate and fraud transactions

```{r}
table(cc$Class)
```

#### Percentage distribution of legitimate and fraud transactions

```{r}
pcnt <- prop.table(table(cc$Class))
pcnt
```

#### Pie chart of credit card transactions

```{r}
labels <- paste(c("legit","fraud"), round((100*pcnt),2),"%")
pie(table(cc$Class),labels,col=c("orange","red"))
```

#### Predictions Without Model And Confusion-Matrix
```{r}
predictions <- rep.int(0, nrow(cc))
predictions <- factor(predictions, levels = c(0,1))

confusionMatrix(data = predictions, reference = cc$Class)
```

### Sampling the data

#### Creating sample data-set

```{r}
set.seed(1)
cc_sample <- cc %>% sample_frac(0.1)
```

### Exploring sampled data

#### Checking structure of sample data-set

```{r}
str(cc_sample)
```

#### Checking distribution of "Class" column

```{r}
table(cc_sample$Class)
```

#### Plotting scatter plot showing distribution of class column 

```{r}
ggplot(data = cc_sample, aes(x = V1, y = V2, col = Class))+
  geom_point()+
  theme_bw()+
  scale_color_manual(values = c("dodgerblue2", "red"))
```

### Splitting the data-set

#### Creating training and test sets for fraud detection 

```{r}
set.seed(1)
data_sample <- sample.split(cc_sample$Class,SplitRatio = 0.80)
train_data <- subset(cc_sample, data_sample==TRUE)
test_data <- subset(cc_sample, data_sample==FALSE)
```

#### Checking structure of train data

```{r}
str(train_data)
```

#### Checking structure of test data

```{r}
str(test_data)
```

#### Checking dimensions of train data

```{r}
dim(train_data)
```

#### Checking dimensions of test data

```{r}
dim(test_data)
```

### Random Over Sampling (ROS) Method

#### Checking distribution of "Class" in train data

```{r}
table(train_data$Class)
```

#### Setting threshold for legit transaction

```{r}
n_legit <- 1594
new_frac_legit <- 0.50
new_n_total <- n_legit/new_frac_legit 
new_n_total

```

#### Performing Random-Over-Sampling (ROS)
```{r}
oversampling_result <- ovun.sample(Class ~ .,
                                   data = train_data,
                                   method = "over",
                                   N = new_n_total,
                                   seed = 2021)
```

#### Storing Random-Over-Sampling (ROS) results

```{r}
oversampled_credit <- oversampling_result$data
```

#### Checking distribution of "Class" in Random-Over-Sampling (ROS) results

```{r}
table(oversampled_credit$Class)
```

#### Plotting distribution of "Class" in Random-Over-Sampling (ROS) results

```{r}
ggplot(data = oversampled_credit, aes(x = V1, y = V2, col = Class))+
  geom_point(position = position_jitter(width = 0.1))+
  theme_bw()+
  scale_color_manual(values = c("dodgerblue2", "red"))
```

### Random Under Sampling (RUS) Method  

#### Checking distribution of "Class" in train data

```{r}
table(train_data$Class)
```

#### Setting threshold for fraud transaction

```{r}
n_fraud <- 6
new_frac_fraud <- 0.50
new_n_total <- n_fraud/new_frac_fraud
new_n_total
```

#### Performing Random-Under-Sampling (RUS)

```{r}
undersampling_result <- ovun.sample(Class ~ .,
                                   data = train_data,
                                   method = "under",
                                   N = new_n_total,
                                   seed = 2021)
```

#### Storing Random-Under-Sampling (RUS) results

```{r}
undersampled_credit <- undersampling_result$data
```

#### Checking distribution of "Class" in Random-Under-Sampling (RUS) results

```{r}
table(undersampled_credit$Class)
```

#### Plotting distribution of "Class" in Random-Under-Sampling (RUS) results

```{r}
ggplot(data = undersampled_credit, aes(x = V1, y = V2, col = Class))+
  geom_point()+#(position = position_jitter(width = 0.1))
  theme_bw()+
  scale_color_manual(values = c("dodgerblue2", "red"))
```

### Using Both Random-Over-Sampling (ROS) And Random-Under-Sampling (RUS) Methods

#### Storing number of rows of train data
```{r}
n_new <- nrow(train_data)
n_new
```

#### Setting threshold

```{r}
fraction_fraud_new <- 0.50
```

#### Performing Random-Over-Sampling (ROS) AND Random-Under-Sampling (RUS)

```{r}
sampling_result <- ovun.sample(Class ~ .,
                               data = train_data,
                               method = "both",
                               N = n_new,
                               p = fraction_fraud_new,
                               seed = 2021)
```

#### Storing Sampling results

```{r}
sampled_credit <- sampling_result$data
```

#### Checking properties of sampling results

```{r}
nrow(sampled_credit)
table(sampled_credit$Class)
prop.table(table(sampled_credit$Class))
```

#### Plotting distribution of "Class" in sampling results

```{r}
ggplot(data = sampled_credit, aes(x = V1, y = V2, col = Class))+
  geom_point()+#(position = position_jitter(width = 0.1))
  theme_bw()+
  scale_color_manual(values = c("dodgerblue2", "red"))
```

### Using SMOTE to balance the data-set

#### Checking distribution of "Class" in train data

```{r}
table(train_data$Class)
```

#### Set no. of fraud & legitimate cases and desired % of legitimate cases

```{r}
n0 <- 1594
n1 <- 6
r0 <- 0.6
```

#### Calculate the value for the dup_size parameter of SMOTE

```{r}
ntimes <- ((1-r0)/r0) * (n0/n1) - 1
```

#### Performing SMOTE operation

```{r}
smote_output <- SMOTE(X = train_data[ ,-c(1,31)],
                      target = train_data$Class,
                      K = 5,
                      dup_size = ntimes)
```

#### Storing SMOTE-Balanced data

```{r}
credit_smote <- smote_output$data
```

#### Changing column-name for column-30 to "Class"

```{r}
colnames(credit_smote)[30] <- "Class"
```

#### Checking properties of SMOTE data

```{r}
nrow(credit_smote)
table(credit_smote$Class)
prop.table(table(credit_smote$Class))
```

#### Plotting distribution of "Class" in sampling results

```{r}
ggplot(data = credit_smote, aes(x = V1, y = V2, col = Class))+
  geom_point()+#(position = position_jitter(width = 0.1))
  theme_bw()+
  scale_color_manual(values = c("dodgerblue2", "red"))
```

## (A) Model Building, Prediction And Accuracy (For Sampled Data)

### (1) Using SMOTE-Balanced Data

### Building Decision-Tree Model using SMOTE-balanced data

```{r}
# Building Model
CART_model <- rpart(Class~., credit_smote)

# Checking Summary
summary(CART_model)
```

#### Plotting decision-tree model 

```{r}
rpart.plot(CART_model, extra = 0, type = 5, tweak = 1.2)
```

#### Predicting fraud cases

```{r}
predicted_val <- predict(CART_model, test_data, type = "class") 

# Viewing Predictions
head(predicted_val)
```

#### Building Confusion-Matrix

```{r}
confusionMatrix(predicted_val, test_data$Class)
```

### (2) Without Using SMOTE-Balanced Data

### Building Decision-Tree Model without using SMOTE-balanced data

```{r}
# Building Model
CART_model <- rpart(Class~., train_data[,-1])

# Checking Summary
```

#### Plotting decision-tree model 

```{r}
rpart.plot(CART_model, extra = 0, type = 5, tweak = 1.2)
```

#### Predicting fraud cases

```{r}
predicted_val <- predict(CART_model, test_data[,-1], type = "class")
head(predicted_val)
```

#### Building Confusion-Matrix

```{r}
confusionMatrix(predicted_val, test_data$Class)
```

## (B) Model Building, Prediction And Accuracy (For Whole Data)

### (1) Using SMOTE-Balanced data

#### Building decision-tree model 

```{r}
CART_model <- rpart(Class~., credit_smote)
summary(CART_model)
```

#### Plotting decision-tree model 

```{r}
rpart.plot(CART_model, extra = 0, type = 5, tweak = 1.2)
```

#### Predicting fraud cases

```{r}
predicted_val <- predict(CART_model, cc[,-1], type = "class") 
head(predicted_val)
```

#### Building Confusion-Matrix

```{r}
confusionMatrix(predicted_val, cc$Class)
```

### (2) Without Using SMOTE-Balanced Data

#### Building decision-tree model without using SMOTE data

```{r}
CART_model <- rpart(Class~., train_data[,-1])
summary(CART_model)
```

#### Plotting decision-tree model 

```{r}
rpart.plot(CART_model, extra = 0, type = 5, tweak = 1.2)
```

#### Predicting fraud cases

```{r}
predicted_val <- predict(CART_model, cc[,-1], type = "class")
head(predicted_val)
```

#### Building Confusion-Matrix

```{r}
confusionMatrix(predicted_val, cc$Class)
```