---
title: "R : Module 7 : Decision Tree"
output:
  html_document:
    df_print: paged
header-includes: \usepackage{fvextra} \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      tidy = 'formatR',
                      #tidy.opts = list(blank = FALSE, width.cutoff = 90),
                      prompt = F,
                      size = 'large')
```

## Module 7 : Assignment 1-6

### Loading required packages

```{r}
library('dplyr')
library('caTools')
library('caret')
library('rpart')
```

### (A) City-Temperature Data-Set

#### Preparing the data

```{r}
# Loading the data-set
city_1 <- read.csv('/home/carb0n/Desktop/r/d/city_temperature.csv')

# Checking for NA values
sum(is.na(city_1))

# Removing "State" column as it is not needed
city_1 <- within(city_1, rm(State))

# Checking structure of data-set
str(city_1)

# Checking unique values in Year column
unique(city_1$Year)

# Replacing unique values "201" and "200" from Year column to "2001" and "2000" respectively 
for(i in 1:nrow(city_1)){
  if(city_1$Year[i]==200){
    city_1$Year[i]=2000
  }
  else if(city_1$Year[i]==201){
    city_1$Year[i]=2001
  }
}

# Converting appropriate variables to factor-type 
as.factor(city_1$Region) -> city_1$Region
as.factor(city_1$Country) -> city_1$Country
as.factor(city_1$City) -> city_1$City
as.factor(city_1$Year) -> city_1$Year
```

#### Splitting the data-set into 70:30 ratio based on dependent feature

```{r}
set.seed(1)
sample.split(city_1$AvgTemperature, SplitRatio=0.70) -> city_1_split_tag
subset(city_1, city_1_split_tag==T) -> city_1_train
subset(city_1, city_1_split_tag==F) -> city_1_test
```

#### Decision-Tree Model for city-temperature data-set

```{r}
# Building Model
city_tree1 <- rpart(Region ~ AvgTemperature, data = city_1_train, method = "class")

# Checking misclassification error rate and summary of model
summary(city_tree1)
```

#### Predicting on test-data

```{r}
result_city_tree1 <- predict(city_tree1, newdata = city_1_test, type = "class")
```

#### Confusion-Matrix 

```{r}
confusionMatrix(city_1_test$Region, result_city_tree1)
```

### (B) Customer-Churn Data-Set

#### Preparing the data

```{r}
# Loading the data-set
churn_4 <- read.csv('/home/carb0n/Desktop/r/d/customer_churn.csv')

# Checking structure of data-set
str(churn_4)

# Checking for NA values
colSums(is.na(churn_4))

# Removing "CustomerID" column as it is not needed
churn_4 <- within(churn_4, rm(customerID))

# Plotting box-plot for TotalCharges column to check for outliers
boxplot(churn_4$TotalCharges)

# Replacing NA values in TotalCharges column with mean value for TotalCharges column
churn_4$TotalCharges[is.na(churn_4$TotalCharges)] <- mean(churn_4$TotalCharges, na.rm = TRUE)

# Converting appropriate variables to factor-type 
as.factor(churn_4$gender) -> churn_4$gender
as.factor(churn_4$SeniorCitizen) -> churn_4$SeniorCitizen
as.factor(churn_4$Partner) -> churn_4$Partner
as.factor(churn_4$Dependents) -> churn_4$Dependents
as.factor(churn_4$PhoneService) -> churn_4$PhoneService
as.factor(churn_4$MultipleLines) -> churn_4$MultipleLines
as.factor(churn_4$InternetService) -> churn_4$InternetService
as.factor(churn_4$OnlineSecurity) -> churn_4$OnlineSecurity
as.factor(churn_4$OnlineBackup) -> churn_4$OnlineBackup
as.factor(churn_4$DeviceProtection) -> churn_4$DeviceProtection
as.factor(churn_4$TechSupport) -> churn_4$TechSupport
as.factor(churn_4$StreamingTV) -> churn_4$StreamingTV
as.factor(churn_4$StreamingMovies) -> churn_4$StreamingMovies
as.factor(churn_4$Contract) -> churn_4$Contract
as.factor(churn_4$PaperlessBilling) -> churn_4$PaperlessBilling
as.factor(churn_4$PaymentMethod) -> churn_4$PaymentMethod
as.factor(churn_4$Churn) -> churn_4$Churn
```

#### Splitting the data-set into 80:20 ratio based on dependent feature

```{r}
set.seed(1)
sample.split(churn_4$Churn, SplitRatio=0.80) -> churn_4_split_tag
subset(churn_4, churn_4_split_tag==T) -> churn_4_train
subset(churn_4, churn_4_split_tag==F) -> churn_4_test
```

#### Decision-Tree Model for customer-churn data-set

```{r}
# Building Model
churn_tree1 <- rpart(Churn ~ MonthlyCharges, data = churn_4_train, method = "class")

# Checking misclassification error rate and summary of model
summary(churn_tree1)
```

#### Predicting on test-data

```{r}
result_churn_tree1 <- predict(churn_tree1, newdata = churn_4_test, type = "class")
```

#### Confusion-Matrix 

```{r}
confusionMatrix(churn_4_test$Dependents, result_churn_tree1)
```

### (C) Pharmacovigilance Data-Set

#### Preparing the data

```{r}
# Loading the data-set
pharma_3 <- read.csv('/home/carb0n/Desktop/r/d/Pharmacovigilance_audit_Data.csv')

# Removing LocationID and PatientID columns as they are not needed
pharma_3 <- within(pharma_3, rm(LocationID, PatientID))

# Checking for NA values
colSums(is.na(pharma_3))

# Checking structure of the data-set
str(pharma_3)

# Converting appropriate variables to factor-type
as.factor(pharma_3$Issues) -> pharma_3$Issues
as.factor(pharma_3$DrugID) -> pharma_3$DrugID
as.factor(pharma_3$Gender) -> pharma_3$Gender
```

#### Splitting the data-set into 75:25 ratio based on dependent feature

```{r}
set.seed(1)
sample.split(pharma_3$DrugID, SplitRatio=0.75) -> pharma_3_split_tag
subset(pharma_3, pharma_3_split_tag==T) -> pharma_3_train
subset(pharma_3, pharma_3_split_tag==F) -> pharma_3_test
```

#### Decision-Tree Model for pharmacovigilance data-set

```{r}
# Building Model
pharma_tree1 <- rpart(DrugID ~ Age, data = pharma_3_train, method = "class")

# Checking misclassification error rate and summary of model
summary(pharma_tree1)
```

#### Predicting on test-data

```{r}
result_pharma_tree1 <- predict(pharma_tree1, newdata = pharma_3_test, type = "class")
```

#### Confusion-Matrix 

```{r}
confusionMatrix(pharma_3_test$DrugID, result_pharma_tree1)
```

## Module 7: Case Study 1 - Decision Tree

### Loading required packages

```{r}
library('dplyr')
library('caTools')
library('caret')
library('rpart')
```

#### Task-1 : Building a decision tree model

```{r}
#a. Start off by dividing the 'customer_churn' data into train & test sets in 70:30 ratios. The split-criteria would be determined by the 'Dependents' column
#b. Build a decision tree model on top of the 'train' set, where the dependent variable is 'Dependents' & the independent variable is 'Partner'. Store the result in 'mod_tree1'
#c. Plot the tree and add text
#d. Predict the values on the test set and store the result in 'result_tree1'
#e. Build a confusion matrix for the actual values & the predicted values
#f. Calculate the accuracy from the confusion matrix
```

#### Preparing the data

```{r}
# Loading the data-set
churn_4 <- read.csv('/home/carb0n/Desktop/r/d/customer_churn.csv')

# Checking structure of data-set
str(churn_4)

# Checking for NA values
colSums(is.na(churn_4))

# Removing "CustomerID" column as it is not needed
churn_4 <- within(churn_4, rm(customerID))

# Plotting box-plot for TotalCharges column to check for outliers
boxplot(churn_4$TotalCharges)

# Replacing NA values in TotalCharges column with mean value for TotalCharges column
churn_4$TotalCharges[is.na(churn_4$TotalCharges)] <- mean(churn_4$TotalCharges, na.rm = TRUE)

# Converting appropriate variables to factor-type 
as.factor(churn_4$gender) -> churn_4$gender
as.factor(churn_4$SeniorCitizen) -> churn_4$SeniorCitizen
as.factor(churn_4$Partner) -> churn_4$Partner
as.factor(churn_4$Dependents) -> churn_4$Dependents
as.factor(churn_4$PhoneService) -> churn_4$PhoneService
as.factor(churn_4$MultipleLines) -> churn_4$MultipleLines
as.factor(churn_4$InternetService) -> churn_4$InternetService
as.factor(churn_4$OnlineSecurity) -> churn_4$OnlineSecurity
as.factor(churn_4$OnlineBackup) -> churn_4$OnlineBackup
as.factor(churn_4$DeviceProtection) -> churn_4$DeviceProtection
as.factor(churn_4$TechSupport) -> churn_4$TechSupport
as.factor(churn_4$StreamingTV) -> churn_4$StreamingTV
as.factor(churn_4$StreamingMovies) -> churn_4$StreamingMovies
as.factor(churn_4$Contract) -> churn_4$Contract
as.factor(churn_4$PaperlessBilling) -> churn_4$PaperlessBilling
as.factor(churn_4$PaymentMethod) -> churn_4$PaymentMethod
as.factor(churn_4$Churn) -> churn_4$Churn
```

#### Splitting the data-set into 70:30 ratio based on dependent feature

```{r}
set.seed(1)
sample.split(churn_4$Dependents, SplitRatio=0.70) -> churn_4_split_tag
subset(churn_4, churn_4_split_tag==T) -> churn_4_train
subset(churn_4, churn_4_split_tag==F) -> churn_4_tests
```

#### Decision-Tree 

```{r}
# Building Model
mod_tree1 <- rpart(Dependents ~ Partner, data = churn_4_train, method = "class")

# Checking misclassification error rate and summary of model
summary(mod_tree1)
```

#### Plotting Decision-Tree

```{r}
# Plotting the model
plot(mod_tree1)

# Giving labels to plotted decision-tree model
text(mod_tree1, pretty = 1)
```

#### Predicting on test-data

```{r}
result_tree1 <- predict(mod_tree1, newdata = churn_4_test, type = "class")
```
 
#### Confusion-Matrix 

```{r}
confusionMatrix(churn_4_test$Dependents, result_tree1)
```

#### Task-2 : Building 2nd decision tree model on same 'train' & 'test' sets 

```{r}
#a. In this case the dependent variable is 'Dependents' & the independent variables are 'Partner' & 'InternetService'. Store the result in 'mod_tree2'
#b. Plot the tree & add text
#c. Predict the values on the test set & store the result in 'result_tree2'
#d. Build a confusion matrix for the actual values & the predicted values
#e. Calculate the accuracy from the confusion matrix
```

#### Decision-Tree

```{r}
# Building Model
mod_tree2 <- rpart(Dependents ~ Partner +InternetService, data = churn_4_train, method = "class")

# Checking misclassification error rate and summary of model
summary(mod_tree2)
```

#### Plotting Decision-Tree

```{r}
plot(mod_tree2)

# Giving labels to plotted decision-tree model
text(mod_tree2, pretty = 1)
```

#### Predicting on test-data

```{r}
result_tree2 <- predict(mod_tree2, newdata = churn_4_test, type = "class")
```

#### Confusion-Matrix 

```{r}
confusionMatrix(churn_4_test$Dependents, result_tree2)
```