---
title: "R : Module 3 : Data Visualisation"
output:
  html_document:
    df_print: paged
header-includes: \usepackage{fvextra} \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      tidy = 'formatR',
                      #tidy.opts = list(blank = FALSE, width.cutoff = 90),
                      prompt = F,
                      size = 'large')
```

## Module 3 : Assignment 1 - AES Attribute

### Loading required packages

```{r}
library(ggplot2)
```

```{r}
# Loading the data-set
customer_churn <- read.csv('/home/carb0n/Desktop/r/d/customer_churn.csv')

# 1. Create the following bar-plots where:

# a. PhoneService column is assigned to the x aesthetic.
ggplot(data=customer_churn,aes(x=PhoneService))+geom_bar()

# b. Gender column is assigned to the x aesthetic.
ggplot(data=customer_churn,aes(x=gender))+geom_bar()

# c. InternetService is assigned to x aesthetic and MonthlyCharges to aesthetic.
ggplot(data=customer_churn,aes(x=InternetService))+
  geom_bar(aes(MonthlyCharges))

# d. MonthlyCharges is assigned to x aesthetic and SeniorCitizen to aesthetic.
ggplot(data=customer_churn,aes(x=MonthlyCharges))+
  geom_bar(aes(SeniorCitizen))

# e. TotalCharges is assigned to x aesthetic and Tenure to aesthetic.
ggplot(data=customer_churn,aes(x=TotalCharges))+
  geom_bar(aes(tenure))

# f. StreamingTV is assigned to x aesthetic and SeniorCitizen to aesthetic.
ggplot(data=customer_churn,aes(x=StreamingTV))+
  geom_bar(aes(SeniorCitizen))

# g. Dependents is assigned to x aesthetic and Partner to aesthetic. 
ggplot(data=customer_churn,aes(x=Dependents))+
  geom_bar(aes(Partner))
```

## Module 3 : Assignment 2 - bins attribute

### Loading required packages

```{r}
library(ggplot2)
```

```{r}
# Loading the data-set
customer_churn <- read.csv('/home/carb0n/Desktop/r/d/customer_churn.csv')

# Loading the data-set
pharma1 <- read.csv('/home/carb0n/Desktop/r/d/Pharmacovigilance_audit_Data.csv')

# 1. Create a histogram for 'Age' column using pharmacovigilance dataset:
# a. Set number of bins to 100.
# b. Assign as color 'azure' to the histogram.
# c. Assign a color 'white' to the 'fill' attribute in geom_histogram function.
ggplot(data=pharma1,aes(x=Age))+
  geom_histogram(bins=100,col='azure',fill='white')

# 2. Create a histogram for 'Patient ID' using pharmacovigilance dataset:
# a. Assign a color 'wheat3' to the plot.
# b. Set number of bins to 50.
# c. Assign a color 'black' to the 'fill' attribute in geom_histogram function.
ggplot(data=pharma1,aes(x=PatientID))+
  geom_histogram(col='wheat3',bins=50,fill='black')

# 3. Create a histogram using customer churn dataset:
# a. Assign 'MonthlyCharges' column to the x-axis.
# b. Set the number of bins to 80.
# c. Assign a color 'violet' to the bars.
# d. Assign a color 'white' to the 'fill' attribute in geom_histogram function.
ggplot(data=customer_churn,aes(x=MonthlyCharges))+
  geom_histogram(bins=80,col='violet',fill='white')

# 4. Create a histogram using customer churn data:
# a. Assign 'tenure' column to the x-axis.
# b. Set the number of bins to 50.
# c. Assign a color 'white' to the bars.
# d. Assign a color 'black' to the 'fill' attribute in geom_histogram function. 
ggplot(data=customer_churn,aes(x=tenure))+
  geom_histogram(bins=50,col='white',fill='black')
```

## Module 3 : Assignment 3 - col attribute

### Loading required packages

```{r}
library(ggplot2)
```

```{r}
# Loading the data-set
placement1 <- read.csv('/home/carb0n/Desktop/r/d/Placement_Data_Full_Class.csv')

# 1.Create the following bar-plot where:

# a.'Gender' is assigned to x-axis, 'ssc_p' to y-axis, and map 'ssc_b' with col aesthetic.
ggplot(data=placement1,aes(x=gender,y=ssc_p,col=ssc_b))+
  geom_bar(stat='identity')

# b.'ssc_b' is assigned to x-axis, 'hsc_s' to y-axis, and map 'degree_t' with col aesthetic.
ggplot(data=placement1,aes(x=ssc_b,y=hsc_s,col=degree_t))+
  geom_bar(stat='identity')

# c.'Status' is assigned to x-axis, salary to y-axis, and map 'specialisation' with col aesthetic.
ggplot(data=placement1,aes(x=status,y=salary,col=specialisation))+
  geom_bar(stat='identity')

# d.'Degree_t' is assigned to x-axis, 'etest_p' to y-axis, and map 'workex' with col aesthetic.
ggplot(data=placement1,aes(x=degree_t,y=etest_p,col=workex))+
  geom_bar(stat='identity')

# e.'Hsc_s' is assigned to x-axis, 'hsc_b' to y-axis, and map gender with col aesthetic.
ggplot(data=placement1,aes(x=hsc_s,y=hsc_b,col=gender))+
  geom_bar(stat='identity')

# f.'Hsc_s' is assigned to x-axis, 'etest_p' to y-axis, and map 'degree_t' with col aesthetic.
ggplot(data=placement1,aes(x=hsc_s,y=etest_p,col=degree_t))+
  geom_bar(stat='identity')
```

## Module 3 : Assignment 4 - facet_grid() 

### Loading required packages

```{r}
library(ggplot2)
```

```{r}
# Loading the data-set
customer_churn <- read.csv('/home/carb0n/Desktop/r/d/customer_churn.csv')

# Loading the data-set
placement1 <- read.csv('/home/carb0n/Desktop/r/d/Placement_Data_Full_Class.csv')

# 1.Build a histogram for the 'TotalCharges' column. Map 'gender' onto fill aesthetic.
# a.Facet the plot w.r.t 'gender' column
ggplot(data=customer_churn,aes(x=TotalCharges,fill=gender))+
  geom_histogram()+facet_grid(~gender)

# 2.Build a box-plot between 'tenure' & 'InternetService'.Map 'tenure' on the y-axis & 'InternetService' on the x-axis. Map 'Contract' to the fill aesthetic
# a.Facet the plot w.r.t 'Contract' column
ggplot(data=customer_churn,aes(y=tenure,x=InternetService,fill=Contract))+
  geom_boxplot()+facet_grid(~Contract)

# 3.Build a scatter-plot between 'etest_p' & 'mba_p'.Map 'etest_p' on the y-axis & 'mba_p' on the x-axis. Map 'degree_t' onto col aesthetic
# a.Facet the plot w.r.t 'degree_t' column
ggplot(data=placement1,aes(y=etest_p,x=mba_p,fill=degree_t))+
  geom_boxplot()+facet_grid(~degree_t)
```

## Module 3 : Assignment 5 - fill attribute

### Loading required packages

```{r}
library(ggplot2)
```

```{r}
# Loading the data-set
placement1 <- read.csv('/home/carb0n/Desktop/r/d/Placement_Data_Full_Class.csv')

# 1.Build a bar-plot:

# a.Assign 'etest_p' column to x-axis
ggplot(data=placement1,aes(x=etest_p))+geom_bar()

# b.Assign 'specialisation'column to x-axis
ggplot(data=placement1,aes(x=specialisation))+geom_bar()

# c.Assign 'workex' to the fill attribute
ggplot(data=placement1,aes(x=specialisation,fill=workex))+geom_bar()


# 2.Build a bar-plot for the 'degree_t' column:
# a.Assign the fill color to be 'orange'
# b.Assign the boundary color to be 'peru'
ggplot(data=placement1,aes(x=degree_t))+
  geom_bar(fill='orange',col='peru')


# 3.Build a bar-plot for the 'hsc_s' column assigned to x-axis and then:

# a.Assign 'hsc_b' to the fill aesthetic
ggplot(data=placement1,aes(x=hsc_s,fill=hsc_b))+geom_bar()

# b.Assign 'degree_t' to the fill aesthetic
ggplot(data=placement1,aes(x=hsc_s,fill=degree_t))+geom_bar()

# c.Change the position of bars to 'identity'
ggplot(data=placement1,aes(x=hsc_s,fill=degree_t))+
  geom_bar(position='identity')
```

## Module 3 : Assignment 6 - geom_bar attribute

### Loading required packages

```{r}
library(ggplot2)
```

```{r}
# Loading the data-set
pharma1 <- read.csv('/home/carb0n/Desktop/r/d/Pharmacovigilance_audit_Data.csv')

# 1.Build a bar-plot:
# a.Assign 'LocationID' column to x-axis.
# b.Assign 'Issues' to the fill attribute.
# c.Set title to 'Location ID Plot'
ggplot(data=pharma1,aes(x=LocationID,fill=Issues))+
  geom_bar()+ggtitle('Location ID Plot')

# 2.Build a bar-plot for the 'DrugID' column:
# a.Assign the fill color to be 'orange'
# b.Assign the boundary color to be 'peru'
# c.Set title to 'Drug ID Plot'
ggplot(data=pharma1,aes(x=DrugID))+
  geom_bar(fill='orange',col='peru')+ggtitle('Drug ID Plot')

# 3.Build a bar-plot for the 'Gender' column assigned to x-axis and then:
# a.Assign 'DrugID' to the fill aesthetic
# b.Assign 'Issues' to the fill aesthetic
# c.Change the position of bars to 'identity'
# d.Set title to 'Gender Plot'
ggplot(data=pharma1,aes(x=Gender,fill=Issues))+
  geom_bar(position='identity')+ggtitle('Gender Plot')
```

## Module 3 : Assignment 7 - geom_point attribute

### Loading required packages

```{r}
library(ggplot2)
```

```{r}
# Loading the data-set
placement1 <- read.csv('/home/carb0n/Desktop/r/d/Placement_Data_Full_Class.csv')

# 1.Create a scatter-plot:

# a.Assign 'hsc_p' column to x-axis and 'ssc_p' column to y-axis.
ggplot(data=placement1,aes(x=hsc_p, y=ssc_p))+geom_point()

# b.Map 'gender' to the col aesthetic.
ggplot(data=placement1,aes(x=hsc_p, y=ssc_p,col=gender))+geom_point()

# c.Map 'workex' to the col aesthetic.
ggplot(data=placement1,aes(x=hsc_p, y=ssc_p,col=workex))+geom_point()

# d.Map 'status' to the col aesthetic.
ggplot(data=placement1,aes(x=hsc_p, y=ssc_p,col=status))+geom_point()

# e.Add a title to the plot as 'Comparing HSC & SSC Percentage'.
ggplot(data=placement1,aes(x=hsc_p, y=ssc_p,col=status))+
  geom_point()+ggtitle('Comparing HSC & SSC Percentage')


# 2.Create a scatter-plot:

# a.Assign 'degree_p' column to the x-axis and 'mba_p' to the y-axis.
ggplot(data=placement1,aes(x=degree_p, y=mba_p))+geom_point()

# b.Use 'col' as an aesthetic and assign it the column 'status'.
ggplot(data=placement1,aes(x=degree_p, y=mba_p, col=status))+geom_point()

# c.Use 'col' as an aesthetic and assign it the column 'specialisation'.
ggplot(data=placement1,aes(x=degree_p, y=mba_p, col=specialisation))+geom_point()

# d.Add a title to the plot as 'Understanding Degree & MBA Percentage'.
ggplot(data=placement1,aes(x=degree_p, y=mba_p, col=specialisation))+
  geom_point()+ggtitle('Understanding Degree & MBA Percentage')


# 3.Create a scatter-plot:

# a.Assign 'etest_p' column to the x-axis and 'salary' to the y-axis.
ggplot(data=placement1,aes(x=etest_p, y=salary))+geom_point()

# b.Use 'col' as an aesthetic and assign it the column 'workex'.
ggplot(data=placement1,aes(x=etest_p, y=salary, col=workex))+geom_point()

# c.Use 'col' as an aesthetic and assign it the column 'gender'.
ggplot(data=placement1,aes(x=etest_p, y=salary, col=gender))+geom_point()

# d.Add a title to the plot as 'E-test & Salary'
ggplot(data=placement1,aes(x=etest_p, y=salary, col=gender))+
  geom_point()+ggtitle('E-test & Salary')
```

## Module 3 : Assignment 8 - theme()

### Loading required packages

```{r}
library(ggplot2)
```

```{r}
# Loading the data-set
placement1 <- read.csv('/home/carb0n/Desktop/r/d/Placement_Data_Full_Class.csv')

# 1.Build a bar-plot for the 'specialisation' column.Give it a fill color of 'skyblue'
# a.Give the panel a background color of 'cornsilk4'.
# b.Give the plot a background color of 'lightgreen'.
ggplot(data=placement1,aes(x=specialisation))+geom_bar(fill='skyblue')+
  theme(panel.background = element_rect(fill='lightgreen'))


# 2.Build a scatter-plot between 'hsc_p' & 'ssc_p'.Map 'hsc_p' on the y-axis & 'ssc_p' on the x-axis. Assign a color of 'lightgreen' to the points.
# a.Add a title to the plot 'HSC Percent vs SSC Percent'.
# b.Give the panel a background color of 'skyblue'.
# c.Give the plot a background color of 'beige'.
# d.Center align the title.
ggplot(data=placement1,aes(x=ssc_p,y=hsc_p))+geom_point(col='lightgreen')+
  ggtitle('HSC Percent vs SSC Percent')+
  theme(panel.background = element_rect(fill='beige'),plot.title = element_text(hjust = 0.5))
```

## Module 3 : Assignment 9 - xlab and ylab attribute

### Loading required packages

```{r}
library(ggplot2)
```

```{r}
# Loading the data-set
placement1 <- read.csv('/home/carb0n/Desktop/r/d/Placement_Data_Full_Class.csv')

# 1.Create a plot using plot() function for hsc_s column, assign a color 'orange', and label x-axis as 'Higher Secondary Subjects'.
plot(as.factor(placement1$hsc_s),col='orange',xlab='Higher Secondary Subjects')


# 2.Create a plot using plot() function for 'degree_t' column:
# a.Assign a color 'palegreen4'
# b.Label x-axis as 'Percentage in Graduation'
# c.Label y-axis as 'Number of Students'
plot(as.factor(placement1$degree_t),col='palegreen4',xlab='Percentage in Graduation',ylab='Number of Students')


# 3.Create a plot using plot() function for 'specialisation' column & Assign a color 'wheat3':
# a.Label x-axis as 'Specialization Field'
# b.Label y-axis as 'Number of Students'
# c.Give a heading using 'main' attribute as 'Distribution of Specialization'
plot(as.factor(placement1$specialisation),col='wheat3',xlab='Specialisation Field',ylab='Number of Students',main='Distribution of Specialization')
```

## Module 3 : Assignment 10 - geom_boxplot() 

### Loading required packages

```{r}
library(ggplot2)
```

```{r}
# Loading the data-set
placement1 <- read.csv('/home/carb0n/Desktop/r/d/Placement_Data_Full_Class.csv')

# 1.Create a box-plot between 'etest_p' & 'workex'.Map 'etest_p' to the y-axis &   'workex' to the 'x-axis'
# a.Assign it a fill color of 'wheat3'
# b.Assign it a boundary color of 'snow3'
# c.Give a title as 'E-test & Workex'
ggplot(data=placement1,aes(x=workex, y=etest_p))+geom_boxplot(fill='wheat3',col='snow3')+ggtitle('E-test & Workex')


# 2.Build a box-plot between 'etest_p' & 'gender'.Map 'etest_p' to the y-axis & 'gender' to the 'x-axis'

# a.Assign 'degree_t' to the fill aesthetic
ggplot(data=placement1,aes(x=gender, y=etest_p,fill=degree_t))+geom_boxplot()

# b.Assign 'hsc_s' to the fill aesthetic
ggplot(data=placement1,aes(x=gender, y=etest_p,fill=hsc_s))+geom_boxplot()

# c.Give a title as 'E-test & Gender'
ggplot(data=placement1,aes(x=gender, y=etest_p,fill=hsc_s))+geom_boxplot()+ggtitle('E-test & Gender')


# 3.Build a box-plot between 'etest_p' & 'specialisation'
ggplot(data=placement1,aes(x=specialisation, y=etest_p))+geom_boxplot()

# a.Assign 'ssc_b' to the fill aesthetic
ggplot(data=placement1,aes(x=specialisation, y=etest_p, fill=ssc_b))+geom_boxplot()

# b.Assign 'hsc_b' to the fill aesthetic
ggplot(data=placement1,aes(x=specialisation, y=etest_p, fill=hsc_b))+geom_boxplot()

# c.Give a title as 'E-test & Specialization'
ggplot(data=placement1,aes(x=specialisation, y=etest_p, fill=hsc_b))+geom_boxplot()+ggtitle('E-test & Specialisation')
```

## Module 3 : Assignment 11 - geom_histogram()

### Loading required packages

```{r}
library(ggplot2)
```

```{r}
# Loading the data-set
placement1 <- read.csv('/home/carb0n/Desktop/r/d/Placement_Data_Full_Class.csv')

# 1.Create a histogram for 'ssc_p' column:
# a.Assign as color 'azure' to the histogram.
# b.Set number of bins to 50.
# c.Assign a color 'cornsilk4' to the 'fill' attribute in geom_histogram function.
# d.Give it a title as 'SSC Percentage'
ggplot(data=placement1,aes(x=ssc_p))+
  geom_histogram(col='azure',bins=50,fill='cornsilk4')+ggtitle('SSC Percentage')

# 2.Create a histogram for 'hsc_p':
# a.Assign a color 'wheat3' to the plot.
# b.Set number of bins to 50.
# c.Assign a color 'black' to the 'fill' attribute in geom_histogram function.
# d.Give it a title as 'HSC Percentage'
ggplot(data=placement1,aes(x=hsc_p))+
  geom_histogram(col='wheat3',bins=50,fill='black')+ggtitle('HSC Percentage')

# 3.Create a histogram as per the following conditions:
# a.Assign 'degree_p' column to the x-axis.
# b.Set the number of bins to 80.
# c.Assign a color 'violet' to the bars.
# d.Assign a color 'white' to the 'fill' attribute in geom_histogram function.
# e.Give it a title as 'Degree Percentage'
ggplot(data=placement1,aes(x=degree_p))+
  geom_histogram(bins=80,col='violet',fill='white')+ggtitle('Degree Percentage')

# 4.Create a histogram as per the following condition:
# a.Assign 'etest_p' column to the x-axis.
# b.Set the number of bins to 100.
# c.Assign a color 'white' to the bars.
# d.Assign a color 'black' to the 'fill' attribute in geom_histogram function.
# e.Give it a title as 'E-test Percentage'
ggplot(data=placement1,aes(x=etest_p))+
  geom_histogram(bins=100,col='white',fill='black')+ggtitle('E-test Percentage')
```

## Module 3 : Assignment 12 - labs()

### Loading required packages

```{r}
library(ggplot2)
```

```{r}
# Loading the data-set
placement1 <- read.csv('/home/carb0n/Desktop/r/d/Placement_Data_Full_Class.csv')

# 1.Create bar-plot as per the following condition:
# a.Assign 'degree_t' to the x-axis.
# b.Give title of the plot as 'Degree Plot'.
# c.Assign a color 'yellowgreen'.
ggplot(data=placement1,aes(x=degree_t))+
  geom_bar(fill='yellowgreen')+labs(title='Degree Plot')

# 2.Create bar-plot as per the following condition:
# a.Assign 'ssc_b' to the x-axis.
# b.Give title of the plot as 'SSC Plot'.
# c.Assign a color 'beige'.
ggplot(data=placement1,aes(x=ssc_b))+
  geom_bar(fill='beige')+labs(title='SSC Plot')

# 3.Create bar-plot as per the following condition:
# a.Assign 'hsc_b' to the x-axis.
# b.Give title of the plot as 'HSC Plot'.
# c.Assign a color 'cornsilk4'.
ggplot(data=placement1,aes(x=hsc_b))+
  geom_bar(fill='cornsilk4')+labs(title='HSC Plot')

# 4.Create bar-plot as per the following condition:
# a.Assign 'specialisation' to the x-axis.
# b.Give title of the plot as 'Specialization Plot'.
# c.Assign a color 'pink'.
ggplot(data=placement1,aes(x=specialisation))+
  geom_bar(fill='pink')+labs(title='Specialization Plot')
```

## Module 3 : Assignment 13 - plot()

### Loading required packages

```{r}
library(ggplot2)
```

```{r}
# Loading required packages
library(ggplot2)

# Loading the data-set
placement1 <- read.csv('/home/carb0n/Desktop/r/d/Placement_Data_Full_Class.csv')

# 1.Create a plot to understand the distribution of degree_t column.
plot(as.factor(placement1$degree_t))

# 2.Create a plot to understand the distribution of hsc_s column.
plot(as.factor(placement1$hsc_s))

# 3.Create a plot for specialization column and give a heading as 'Specialization of Candidate'.
plot(as.factor(placement1$specialisation),main='Specialization of Candidate')

# 4.Create a plot using ssc_b, give it a color of 'aquamarine4', and give a heading as 'Type of Board'. 
plot(as.factor(placement1$ssc_b),col='aquamarine4',main='Type of Board')
```

## Module 3 : Assignment 14 - plotly package

### Loading required packages

```{r}
library(plotly)
```

```{r}
# Loading the data-set
pharma1 <- read.csv('/home/carb0n/Desktop/r/d/Pharmacovigilance_audit_Data.csv')

# 1.Build a histogram for the 'Age' column.
plot_ly(data=pharma1,x=~Age,type='histogram')

# a.Color should be determined by 'Issues' column
plot_ly(data=pharma1,x=~Age,type='histogram',color=~Issues)

# b.Color should be determined by 'Gender' column
plot_ly(data=pharma1,x=~Age,type='histogram',color=~Gender)

# 2.Create a scatter plot between 'Age' & 'Gender' as per the following condition:
# a.Map 'Age' on the y-axis
# b.Map 'Gender' on the 'x-axis'
# c.Color should be determined by 'Issues' column
plot_ly(data=pharma1, y=~Age, x=~Gender, type='scatter', mode='markers', color=~Issues)

# 3.Create a box-plot between 'DrugID' & 'Age'.
# a.Map 'DrugID' on the x-axis
# b.Map 'Age' on the y-axis
# c.Color should be determined by 'Gender' column 
plot_ly(data=pharma1, y=~Age, x=~DrugID, type='box', color=~Gender)
```

## Module 3 : Case Study 1 - base graphics

### Loading required packages

```{r}
library(ggplot2)
```

```{r}
# Loading the data-set
customer_churn <- read.csv('/home/carb0n/Desktop/r/d/customer_churn.csv')

# 1.Create a bar-plot for the 'PaymentMethod' column.
as.factor(customer_churn$PaymentMethod) -> customer_churn$PaymentMethod
plot(customer_churn$PaymentMethod)

# a.Assign the color 'burlywood4'
plot(customer_churn$PaymentMethod,col='burlywood4')

# b.Assign the x-axis label to be "Payment Method"
plot(customer_churn$PaymentMethod,col='burlywood4', xlab='Payment Method')

# c.Assign the title to be "Bar-Plot for Payment Method"
plot(customer_churn$PaymentMethod,col='burlywood4', xlab='Payment Method', main='Bar-Plot for Payment Method')

# 2.Create a histogram for the 'TotalCharges' column
# a.Assign the color 'forestgreen'
# b.Assign the x-axis label to be "Total Charges"
# c.Assign the title to be "Histogram for Total Charges"
hist(customer_churn$TotalCharges,col='forestgreen',xlab='Total Charges',main='Histogram for Total Charges')

# 3.Create a density plot for the 'TotalCharges' column
# a.Assign the color 'maroon'
# b.Assign the x-axis label to be "Total Charges"
# c.Assign the title to be "Density plot for Total Charges"
plot(density(customer_churn$TotalCharges,na.rm=T),col='maroon',xlab ='Total Charges',main='Density plot for Total Charges')
```

## Module 3 : Case Study 2 - geom_bar()

### Loading required packages

```{r}
library(ggplot2)
```

```{r}
# Loading the data-set
customer_churn <- read.csv('/home/carb0n/Desktop/r/d/customer_churn.csv')

# 1.Build a bar-plot for the 'PhoneService' column
# a.Assign the fill color to be 'pink'
# b.Assign the boundary color to be 'peru'
ggplot(data = customer_churn,aes(x=PhoneService))+geom_bar(fill='pink',col='peru')


# 2.Build a bar-plot for the 'InternetService' column

# a.Assign 'InternetService' to the fill aesthetic
ggplot(data = customer_churn,aes(x=InternetService,fill=InternetService))+geom_bar()

# b.Assign 'Contract' to the fill aesthetic
ggplot(data = customer_churn,aes(x=InternetService,fill=Contract))+geom_bar()

# c.Change the position of bars to 'identity'
ggplot(data = customer_churn,aes(x=InternetService,fill=Contract))+geom_bar(position='identity')


# 3.Build a bar-plot for 'TechSupport' column
# a.Assign 'Churn' to the fill aesthetic 
ggplot(data = customer_churn,aes(x=TechSupport,fill=Churn))+geom_bar()
```

## Module 3 : Case Study 3 - geom_point()

### Loading required packages

```{r}
library(ggplot2)
```

```{r}
# Loading the data-set
customer_churn <- read.csv('/home/carb0n/Desktop/r/d/customer_churn.csv')

# 1.Build a scatter-plot between 'TotalCharges' & 'tenure'.Map 'TotalCharges' to the y-axis & 'tenure' to the 'x-axis'
# a.Assign it the color 'wheat3'
# b.Use 'col' as an aesthetic and Map 'PaymentMethod' to col
ggplot(data=customer_churn,aes(x=tenure,y=TotalCharges,col=PaymentMethod))+geom_point()

# c.Use 'col' as an aesthetic and Map 'gender' to col
ggplot(data=customer_churn,aes(x=tenure,y=TotalCharges,col=gender))+geom_point()

# d.Map 'Dependents' to both 'col' & 'shape' aesthetics
ggplot(data=customer_churn,aes(x=tenure,y=TotalCharges,col=Dependents,shape=Dependents))+geom_point()


# 2.Build a scatter-plot between 'tenure' & 'MonthlyCharges'.Map 'tenure' to the y-axis & 'MonthlyCharges' to the 'x-axis'
# a.Assign it the color 'yellowgreen'
# b.Use 'col' as an aesthetic and Map 'InternetService' to col
ggplot(data=customer_churn,aes(x=MonthlyCharges,y=tenure,col=InternetService))+geom_point()

# c.Use 'col' as an aesthetic and Map 'Contract' to col 
ggplot(data=customer_churn,aes(x=MonthlyCharges,y=tenure,col=Contract))+geom_point()
```

## Module 3 : Case Study 4 - geom_histogram()

### Loading required packages

```{r}
library(ggplot2)
```

```{r}
# Loading the data-set
customer_churn <- read.csv('/home/carb0n/Desktop/r/d/customer_churn.csv')

# 1.Build a histogram for the 'tenure' column
# a.Assign the fill color to be 'mediumspringgreen'
# b.Assign the boundary color to be 'azure'
# c.Change the number of bins to be 100
ggplot(data=customer_churn,aes(x=tenure))+geom_histogram(fill='mediumspringgreen',color='azure',bins=100)

# 2.Build histogram for the 'MonthlyCharges' column

# a.Assign 'PaymentMethod' to the fill aesthetic
ggplot(data=customer_churn,aes(x=MonthlyCharges,fill=PaymentMethod))+geom_histogram()

# b.Assign 'OnlineBackup' to the fill aesthetic
ggplot(data=customer_churn,aes(x=MonthlyCharges,fill=OnlineBackup))+geom_histogram()

# 3.Build histogram for the 'TotalCharges' column

# a.Assign 'gender' to the fill aesthetic
ggplot(data=customer_churn,aes(x=TotalCharges,fill=gender))+geom_histogram()

# b.Assign 'InternetService' to the fill aesthetic 
ggplot(data=customer_churn,aes(x=TotalCharges,fill=InternetService))+geom_histogram()
```