---
title: "R : Module 6 : Logistic Regression"
output:
  html_document:
    df_print: paged
header-includes: \usepackage{fvextra} \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      tidy = 'formatR',
                      #tidy.opts = list(blank = FALSE, width.cutoff = 90),
                      prompt = F,
                      size = 'large')
```

## Module 6 : Assignment 1-6

### Loading required packages

```{r}
library('dplyr')
library('caTools')
```

### (A) City-Temperature Data-Set

#### Preparing the data

```{r}
# Loading the data-set
city_1 <- read.csv('/home/carb0n/Desktop/r/d/city_temperature.csv')

# Checking for NA values
sum(is.na(city_1))

# Removing "State" column as it is not needed
city_1 <- within(city_1, rm(State))

# Checking structure of data-set
str(city_1)

# Checking unique values in Year column
unique(city_1$Year)

# Replacing unique values "201" and "200" from Year column to "2001" and "2000" respectively 
for(i in 1:nrow(city_1)){
  if(city_1$Year[i]==200){
    city_1$Year[i]=2000
  }
  else if(city_1$Year[i]==201){
    city_1$Year[i]=2001
  }
}

# Converting appropriate variables to factor-type 
as.factor(city_1$Region) -> city_1$Region
as.factor(city_1$Country) -> city_1$Country
as.factor(city_1$City) -> city_1$City
as.factor(city_1$Year) -> city_1$Year
```

#### Splitting the data-set into 70:30 ratio based on dependent feature

```{r}
set.seed(1)
sample.split(city_1$AvgTemperature, SplitRatio=0.70) -> city_1_split_tag
subset(city_1, city_1_split_tag==T) -> city_1_train
subset(city_1, city_1_split_tag==F) -> city_1_test
```

#### Simple Logistic Regression Model for city-temperature data-set

```{r}
# Building Model
glm(Region~AvgTemperature, data = city_1_train, family = 'binomial') -> city_1_model_1

# Checking significance of features, null & residual deviance and summary for the model
summary(city_1_model_1)
```

#### Multiple Logistic Regression Model for city-temperature data-set

```{r}
glm(Region~AvgTemperature +Day +Month +Year, data = city_1_train, family = 'binomial') -> city_1_model_2

# Checking significance of features, null & residual deviance and summary for the model
summary(city_1_model_2)
```

### (B) Customer-Churn Data-Set

#### Preparing the data

```{r}
# Loading the data-set
churn_4 <- read.csv('/home/carb0n/Desktop/r/d/customer_churn.csv')

# Checking structure of data-set
str(churn_4)

# Checking for NA values
colSums(is.na(churn_4))

# Removing "CustomerID" column as it is not needed
churn_4 <- within(churn_4, rm(customerID))

# Plotting box-plot for TotalCharges column to check for outliers
boxplot(churn_4$TotalCharges)

# Replacing NA values in TotalCharges column with mean value for TotalCharges column
churn_4$TotalCharges[is.na(churn_4$TotalCharges)] <- mean(churn_4$TotalCharges, na.rm = TRUE)

# Converting appropriate variables to factor-type 
as.factor(churn_4$gender) -> churn_4$gender
as.factor(churn_4$SeniorCitizen) -> churn_4$SeniorCitizen
as.factor(churn_4$Partner) -> churn_4$Partner
as.factor(churn_4$Dependents) -> churn_4$Dependents
as.factor(churn_4$PhoneService) -> churn_4$PhoneService
as.factor(churn_4$MultipleLines) -> churn_4$MultipleLines
as.factor(churn_4$InternetService) -> churn_4$InternetService
as.factor(churn_4$OnlineSecurity) -> churn_4$OnlineSecurity
as.factor(churn_4$OnlineBackup) -> churn_4$OnlineBackup
as.factor(churn_4$DeviceProtection) -> churn_4$DeviceProtection
as.factor(churn_4$TechSupport) -> churn_4$TechSupport
as.factor(churn_4$StreamingTV) -> churn_4$StreamingTV
as.factor(churn_4$StreamingMovies) -> churn_4$StreamingMovies
as.factor(churn_4$Contract) -> churn_4$Contract
as.factor(churn_4$PaperlessBilling) -> churn_4$PaperlessBilling
as.factor(churn_4$PaymentMethod) -> churn_4$PaymentMethod
as.factor(churn_4$Churn) -> churn_4$Churn
```

#### Splitting the data-set into 80:20 ratio based on dependent feature

```{r}
set.seed(1)
sample.split(churn_4$Churn, SplitRatio=0.80) -> churn_4_split_tag
subset(churn_4, churn_4_split_tag==T) -> churn_4_train
subset(churn_4, churn_4_split_tag==F) -> churn_4_test
```

#### Simple Logistic Regression Model for customer-churn data-set

```{r}
# Building Model
glm(Churn~MonthlyCharges, data = churn_4_train, family = 'binomial') -> churn_4_model_1

# Checking significance of features, null & residual deviance and summary for the model
summary(churn_4_model_1)
```

#### Multiple Logistic Regression Model for customer-churn data-set

```{r}
# Building Model
glm(Churn~MonthlyCharges +tenure +StreamingTV +MonthlyCharges +TotalCharges, data = churn_4_train, family = 'binomial') -> churn_4_model_2

# Checking significance of features, null & residual deviance and summary for the model
summary(churn_4_model_2)
```

### (C) Pharmacovigilance Data-Set

#### Preparing the data

```{r}
# Loading the data-set
pharma_3 <- read.csv('/home/carb0n/Desktop/r/d/Pharmacovigilance_audit_Data.csv')

# Removing LocationID and PatientID columns as they are not needed
pharma_3 <- within(pharma_3, rm(LocationID))

# Checking for NA values
colSums(is.na(pharma_3))

# Checking structure of the data-set
str(pharma_3)

# Converting appropriate variables to factor-type
as.factor(pharma_3$Issues) -> pharma_3$Issues
as.factor(pharma_3$DrugID) -> pharma_3$DrugID
as.factor(pharma_3$Gender) -> pharma_3$Gender
```

#### Splitting the data-set into 75:25 ratio based on dependent feature

```{r}
set.seed(1)

sample.split(pharma_3$DrugID, SplitRatio=0.75) -> pharma_3_split_tag
subset(pharma_3, pharma_3_split_tag==T) -> pharma_3_train
subset(pharma_3, pharma_3_split_tag==F) -> pharma_3_test

nrow(pharma_3_train)
nrow(pharma_3_test)
```

#### Simple Logistic Regression Model for pharmacovigilance data-set

```{r}
# Building Model
glm(DrugID~Age, data = pharma_3_train, family = 'binomial') -> pharma_3_model_1

# Checking significance of features, null & residual deviance and summary for the model
summary(pharma_3_model_1)
```

#### Multiple Logistic Regression Model for pharmacovigilance data-set

```{r}
# Building Model
glm(DrugID~Age +Gender +PatientID, data = pharma_3_train, family = 'binomial') -> pharma_3_model_2

# Checking significance of features, null & residual deviance and summary for the model
summary(pharma_3_model_2)
```

### Module 6: Case Study 1 - Multiple Logistic Regression

#### Loading required packages

```{r}
library('dplyr')
library('caTools')
library('caret')
```

#### Task-1 : Build a multiple logistic regression model

```{r}
#a. Start off by dividing the data-set into 'train' & 'test' sets in 65:35 ratio, with the split-criteria being determined by 'gender' column.
#b. Build a logistic regression model on the train set where the dependent variable is 'gender' & the independent variables are 'Dependents', 'InternetService' & 'Contract' & store the result in 'log_mod_multi'.
#c. Predict the values on top of the test set & store the result in 'result_log_multi'.
#d. Have a look at the range of 'result_log_multi' & build a confusion matrix where the threshold of predicted values is greater than '0.49'
#e. Calculate the accuracy of the model from the confusion matrix.
```

####  Preparing the data

```{r}
# Loading the data-set
churn_4 <- read.csv('/home/carb0n/Desktop/r/d/customer_churn.csv')

# Checking structure of data-set
str(churn_4)

# Checking for NA values
colSums(is.na(churn_4))

# Removing "CustomerID" column as it is not needed
churn_4 <- within(churn_4, rm(customerID))

# Plotting box-plot for TotalCharges column to check for outliers
boxplot(churn_4$TotalCharges)

# Replacing NA values in TotalCharges column with mean value for TotalCharges column
churn_4$TotalCharges[is.na(churn_4$TotalCharges)] <- mean(churn_4$TotalCharges, na.rm = TRUE)

# Converting appropriate variables to factor-type 
as.factor(churn_4$gender) -> churn_4$gender
as.factor(churn_4$SeniorCitizen) -> churn_4$SeniorCitizen
as.factor(churn_4$Partner) -> churn_4$Partner
as.factor(churn_4$Dependents) -> churn_4$Dependents
as.factor(churn_4$PhoneService) -> churn_4$PhoneService
as.factor(churn_4$MultipleLines) -> churn_4$MultipleLines
as.factor(churn_4$InternetService) -> churn_4$InternetService
as.factor(churn_4$OnlineSecurity) -> churn_4$OnlineSecurity
as.factor(churn_4$OnlineBackup) -> churn_4$OnlineBackup
as.factor(churn_4$DeviceProtection) -> churn_4$DeviceProtection
as.factor(churn_4$TechSupport) -> churn_4$TechSupport
as.factor(churn_4$StreamingTV) -> churn_4$StreamingTV
as.factor(churn_4$StreamingMovies) -> churn_4$StreamingMovies
as.factor(churn_4$Contract) -> churn_4$Contract
as.factor(churn_4$PaperlessBilling) -> churn_4$PaperlessBilling
as.factor(churn_4$PaymentMethod) -> churn_4$PaymentMethod
as.factor(churn_4$Churn) -> churn_4$Churn
```

#### Splitting the data-set into 80:20 ratio based on dependent feature

```{r}
set.seed(1)
sample.split(churn_4$Churn, SplitRatio=0.80) -> churn_4_split_tag
subset(churn_4, churn_4_split_tag==T) -> churn_4_train
subset(churn_4, churn_4_split_tag==F) -> churn_4_test
```

#### Multiple Logistic Regression Model

```{r}
# Building Model
glm(gender ~ Dependents +InternetService +Contract, data = churn_4_train, family = 'binomial') -> churn_4_model_2

# Checking significance of features, null & residual deviance and summary for the model
summary(churn_4_model_2)

# Predicting for test data
result_log_multi <- predict(churn_4_model_2, newdata = churn_4_test, type = 'response')

# Checking significance of features, null & residual deviance and summary for the model
summary(result_log_multi)

# Checking range of the values for the model
range(result_log_multi)
```

#### Creating confusion matrix with a threshold

```{r}
pv_model2 <- ifelse(result_log_multi>0.51, "Male", "Female")
pv_model2_f <- as.factor(pv_model2)
head(pv_model2_f)
confusionMatrix(churn_4_test$gender,pv_model2_f)
```

#### Task-2 : Build second logistic regression model on the same 'train' & 'test' sets

```{r}
#a. In this case dependent variable is 'gender' & the independent variables are 'tenure', 'MonthlyCharges' & 'PaymentMethod
#b. Predict the values on top of the test set & store the result in 'result_log_multi2'
#c. Have a look at the range of 'result_log_multi2' & build a confusion matrix where the threshold of predicted values is greater than 0.49
#d. Calculate the accuracy of the model from the confusion matrix
```

#### Multiple Logistic Regression Model

```{r}
# Building Model
churn_4_model_4 <- glm(gender ~ tenure +MonthlyCharges +PaymentMethod, data = churn_4_train, family = "binomial")

# Checking significance of features, null & residual deviance and summary for the model
summary(churn_4_model_4)

# Predicting for test data
result_log_multi2 <- predict(churn_4_model_4, newdata = churn_4_test, type = "response")

# Checking significance of features, null & residual deviance and summary for the model
summary(result_log_multi2)


# Checking range of the values for the model
range(result_log_multi2)
```

#### Creating Confusion with a threshold

```{r}
pv_model3 <- ifelse(result_log_multi2>0.51, "Male", "Female")
pv_model3_f <- as.factor(pv_model3)
head(pv_model3_f)
confusionMatrix(churn_4_test$gender,pv_model3_f)
```

## Module 6: Case Study 2 - ROC Curve

### Loading required packages

```{r}
library('dplyr')
library('caTools')
library('ROCR')
```

#### Task-1 : Building a logistic regression model

```{r}
#a. Start off by dividing the data-set into 'train' & 'test' sets in 80:20 ratio, with the split criteria being determined by 'Churn' column
#b. Build a logistic regression model on the train set where the dependent variable is 'Churn' & the independent variables are 'MonthlyCharges', 'tenure' & 'TechSupport' & store the result in 'log_mod_roc'
#c. Predict the values on top of the test set & store the result in 'result_log_roc'
#d. Use the performance () function from the ROCR package & build the 'Accuracy vs cutoff' plot i.e. Plot the 'ROC' curve
#e. Find out the "area under the curve"
```

#### Preparing the data

```{r}
# Loading the data-set
churn_4 <- read.csv('/home/carb0n/Desktop/r/d/customer_churn.csv')

# Checking structure of data-set
str(churn_4)

# Checking for NA values
colSums(is.na(churn_4))

# Removing "CustomerID" column as it is not needed
churn_4 <- within(churn_4, rm(customerID))

# Plotting box-plot for TotalCharges column to check for outliers
boxplot(churn_4$TotalCharges)

# Replacing NA values in TotalCharges column with mean value for TotalCharges column
churn_4$TotalCharges[is.na(churn_4$TotalCharges)] <- mean(churn_4$TotalCharges, na.rm = TRUE)

# Converting appropriate variables to factor-type 
as.factor(churn_4$gender) -> churn_4$gender
as.factor(churn_4$SeniorCitizen) -> churn_4$SeniorCitizen
as.factor(churn_4$Partner) -> churn_4$Partner
as.factor(churn_4$Dependents) -> churn_4$Dependents
as.factor(churn_4$PhoneService) -> churn_4$PhoneService
as.factor(churn_4$MultipleLines) -> churn_4$MultipleLines
as.factor(churn_4$InternetService) -> churn_4$InternetService
as.factor(churn_4$OnlineSecurity) -> churn_4$OnlineSecurity
as.factor(churn_4$OnlineBackup) -> churn_4$OnlineBackup
as.factor(churn_4$DeviceProtection) -> churn_4$DeviceProtection
as.factor(churn_4$TechSupport) -> churn_4$TechSupport
as.factor(churn_4$StreamingTV) -> churn_4$StreamingTV
as.factor(churn_4$StreamingMovies) -> churn_4$StreamingMovies
as.factor(churn_4$Contract) -> churn_4$Contract
as.factor(churn_4$PaperlessBilling) -> churn_4$PaperlessBilling
as.factor(churn_4$PaymentMethod) -> churn_4$PaymentMethod
as.factor(churn_4$Churn) -> churn_4$Churn
```

#### Splitting the data-set into 80:20 ratio based on dependent feature

```{r}
set.seed(1)
sample.split(churn_4$Churn, SplitRatio=0.80) -> churn_4_split_tag
subset(churn_4, churn_4_split_tag==T) -> churn_4_train
subset(churn_4, churn_4_split_tag==F) -> churn_4_test
```

#### Multiple Logistic Regression Model

```{r}
# Building Model
log_mod_roc <- glm(Churn ~ +MonthlyCharges +tenure +TechSupport, data = churn_4_train, family = "binomial")

# Checking significance of features, null & residual deviance and summary for the model
summary(log_mod_roc)

# Predicting for test data
result_log_roc <- predict(log_mod_roc, newdata = churn_4_test, type = "response")

# Checking significance of features, null & residual deviance and summary for the model
summary(result_log_roc)
```

#### Plotting "Accuracy V/s Cutoff" (ROC Curve)

```{r}
predict_log <- prediction(result_log_roc, churn_4_test$Churn)
performance(predict_log, "acc") -> acc
plot(acc)

roc_curve <- performance(predict_log,"tpr","fpr") 
plot(roc_curve, colorize = T)
```

#### Calculating Area Under the Curve (AUC)

```{r}
auc.tmp <- performance(predict_log,"auc")
auc <- as.numeric(auc.tmp@y.values)
auc
```