---
title: "R : Module 11 : Time Series Analysis"
output:
  html_document:
    df_print: paged
header-includes: \usepackage{fvextra} \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      tidy = 'formatR',
                      #tidy.opts = list(blank = FALSE, width.cutoff = 90),
                      prompt = F,
                      size = 'large')
```

## Module 11 : Case Study 1 - Time Series

### Loading required packages

```{r}
library('aTSA')
library('fUnitRoots')
```

### Exploring the data

```{r}
# Loading the data-set
data("AirPassengers")
ts <- AirPassengers

# Checking class of data
class(ts)

# Checking beginning year of time-series
start(ts)

# Checking end year of time-series
end(ts)

# Checking summary of time-series
summary(ts)

# Viewing cycle of time-series
cycle(ts)
```

### Plotting the data

```{r}
# Plotting time-series
plot(ts)

# Plotting box-plot of number-of-air-passengers (per-year) and cycle of time-series
boxplot(ts~cycle(ts))

#Decomposing the data into its trend, seasonal, and random error components
ts_decom <- decompose(ts, type = "multiplicative")
plot(ts_decom)
```

### Testing Assumptions 

#### Augmented Dickey-Fuller Test (For Stationarity)

```{r}
adfTest(ts)
adfTest(log(ts))
adfTest(diff(ts))
```

#### KPSS-Test (For Stationarity)

```{r}
kpss.test(ts)
```

#### ACF-Test (Auto-Correlation Test)

```{r}
acf(ts)
acf(diff(ts))
```

#### PACF-Test (Partial Auto-Correlation Test)

```{r}
pacf(diff(ts))
plot(diff(log(ts)))
```

##### Based on results from above tests, the time-series satisfies stationarity assumption.

### Building ARIMA model

```{r}
# Creating ARIMA model on "AirPassengers" data-set
mod_time <- arima(log(ts), c(0,1,1), seasonal=list(order=c(0,1,1), period=12))

# Checking model details
mod_time
```

### Predicting for next 20 years

```{r}
# Predicting for next 20 years
pred_time <- predict(mod_time, n.ahead=20*12)
head(pred_time)

# Finding inverse of log for predicted values
pred_time_1 <- round(2.718^pred_time$pred,0)
head(pred_time_1)
```

### Plotting actual values and predicted values (using log scale and dotted lines for predicted values)

```{r}
ts.plot(ts, pred_time_1, log="y", lty=c(1,3))
```