---
title: "R : Module 5 : Linear Regression"
output:
  html_document:
    df_print: paged
header-includes: \usepackage{fvextra} \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      tidy = 'formatR',
                      #tidy.opts = list(blank = FALSE, width.cutoff = 90),
                      prompt = F,
                      size = 'large')
```

## Module 5 : Assignment 1-11

### Loading required packages

```{r}
library('dplyr')
library('caTools')
library('ggplot2')
```

### (A) City-Temperature data-set

#### Preparing the data

```{r}
# Loading the data-set
city_1 <- read.csv('/home/carb0n/Desktop/r/d/city_temperature.csv')

# Checking for NA values
sum(is.na(city_1))

# Removing State column as it is not needed
city_1 <- within(city_1, rm(State))

# Checking structure of data-set
str(city_1)

# Checking unique values in Year column
unique(city_1$Year)

# Replacing unique values "201" and "200" from Year column to "2001" and "2000" respectively 
for(i in 1:nrow(city_1)){
  if(city_1$Year[i]==200){
    city_1$Year[i]=2000
  }
  else if(city_1$Year[i]==201){
    city_1$Year[i]=2001
  }
}

# Converting appropriate variables to factor-type 
as.factor(city_1$Region) -> city_1$Region
as.factor(city_1$Country) -> city_1$Country
as.factor(city_1$City) -> city_1$City
as.factor(city_1$Year) -> city_1$Year
```

#### Splitting the data-set into 70:30 ratio based on dependent feature


```{r}
set.seed(1)
sample.split(city_1$AvgTemperature, SplitRatio=0.70) -> city_1_split_tag
subset(city_1, city_1_split_tag==T) -> city_1_train
subset(city_1, city_1_split_tag==F) -> city_1_test
```

#### Simple Linear Regression Model  for city_temperature data-set
 
```{r}
# Building Model
lm(Month~AvgTemperature, data = city_1_train) -> city_1_model_1

# Checking summary of simple linear regression model (F-Statistics and Adjusted R-Squared)
summary(city_1_model_1)

# Predicting for test data
predict(city_1_model_1, newdata = city_1_test) -> s_pv_1
head(s_pv_1)

# Finding value of error for predictions
cbind(Acutal = city_1_test$Month, Predicted = s_pv_1) -> s_result_1
as.data.frame(s_result_1) -> s_result_1

s_result_1$Acutal - s_result_1$Predicted -> s_error_1
cbind(s_result_1,s_error_1) -> s_result_1
head(s_result_1)

# Finding root mean squared error (RMSE)
sqrt(mean((s_result_1$s_error_1)^2)) -> s_rmse_1
s_rmse_1
```

#### Assumptions for Simple Linear Regression Model for city_temperature data-set

```{r}
# Checking for Linearity
ggplot(data = city_1, aes(x=Month, y=AvgTemperature))+geom_point()+geom_smooth(method = "lm")

# Checking for Equal Error Variance
#ggplot(data = s_result_1, aes(x=Predicted, y=s_error_3))+geom_point()
plot(city_1_model_1, which = 1)

# Checking for Normality of Errors
#qqnorm(s_result_1$s_error_1)
#qqline(s_result_1$s_error_1)
plot(city_1_model_1, which = 2)
```

#### Multiple Linear Regression Model for city_temperature data-set

```{r}
# Building Model
lm(Month~Country +Region +City +AvgTemperature, data = city_1_train) -> city_1_model_2

# Checking summary of multiple linear regression model (F-Statistics and Adjusted R-Squared)
summary(city_1_model_2)

# Predicting for test data
predict(city_1_model_2, newdata = city_1_test) -> m_pv_1
head(m_pv_1)

# Finding value of error for predictions
cbind(Acutal = city_1_test$Month, Predicted = m_pv_1) -> m_result_1
as.data.frame(m_result_1) -> m_result_1

m_result_1$Acutal - m_result_1$Predicted -> m_error_1
cbind(m_result_1,m_error_1) -> m_result_1
head(m_result_1)

# Finding root mean squared error (RMSE)
sqrt(mean((m_result_1$m_error_1)^2)) -> m_rmse_1
m_rmse_1
```

#### Assumptions for Multiple Linear Regression Model for city_temperature data-set

```{r}
# Checking for Linearity
ggplot(data = city_1, aes(x=Month, y=Country))+geom_point()+geom_smooth(method = "lm")
ggplot(data = city_1, aes(x=Month, y=Region))+geom_point()+geom_smooth(method = "lm")
ggplot(data = city_1, aes(x=Month, y=City))+geom_point()+geom_smooth(method = "lm")
ggplot(data = city_1, aes(x=Month, y=AvgTemperature))+geom_point()+geom_smooth(method = "lm")

# Checking for Equal Error Variance
#ggplot(data = m_result_1, aes(x=Predicted, y=m_error_1))+geom_point()
plot(city_1_model_2, which = 1)

# Checking for Normality of Errors
#qqnorm(m_result_1$m_error_1)
#qqline(m_result_1$m_error_1)
plot(city_1_model_2, which = 2)
```

### (B) Placement data-set

#### Preparing the data

```{r}
# Loading the data-set
placement_2 <- read.csv('/home/carb0n/Desktop/r/d/Placement_Data_Full_Class.csv')

# Removing "sl_no" column from data-set as it is not needed
placement_2 <- within(placement_2, rm(sl_no))

# Checking for NA values
colSums(is.na(placement_2))

# Replacing NA values found in salary column with median value of salary column
placement_2$salary[is.na(placement_2$salary)] <- median(placement_2$salary,na.rm=TRUE)

# Checking structure of data-set
str(placement_2)

# Converting appropriate variables to factor-type
as.factor(placement_2$gender) -> placement_2$gender
as.factor(placement_2$ssc_b) -> placement_2$ssc_b
as.factor(placement_2$hsc_b) -> placement_2$hsc_b
as.factor(placement_2$hsc_s) -> placement_2$hsc_s
as.factor(placement_2$degree_t) -> placement_2$degree_t
as.factor(placement_2$workex) -> placement_2$workex
as.factor(placement_2$specialisation) -> placement_2$specialisation
as.factor(placement_2$status) -> placement_2$status
as.integer(placement_2$salary) -> placement_2$salary
```

#### Splitting the data-set into 80:20 ratio based on dependent feature

```{r}
set.seed(1)

sample.split(placement_2$status, SplitRatio=0.80) -> placement_2_split_tag
subset(placement_2, placement_2_split_tag==T) -> placement_2_train
subset(placement_2, placement_2_split_tag==F) -> placement_2_test

nrow(placement_2_train)
nrow(placement_2_test)
```

#### Simple Linear Regression Model  for placement data-set

```{r}
# Building Model
lm(etest_p~mba_p, data = placement_2_train) -> placement_2_model_1

# Checking summary of simple linear regression model (F-Statistics and Adjusted R-Squared)
summary(placement_2_model_1)

# Predicting for test data
predict(placement_2_model_1, newdata = placement_2_test) -> s_pv_2
head(s_pv_2)

# Finding value of error for predictions
cbind(Actual = placement_2_test$etest_p, Predicted = s_pv_2) -> s_result_2
as.data.frame(s_result_2) -> s_result_2

s_result_2$Actual - s_result_2$Predicted -> s_error_2
cbind(s_result_2, s_error_2) -> s_result_2
head(s_result_2)

# Finding root mean squared error (RMSE)
sqrt(mean((s_result_2$s_error_2)^2)) -> s_rmse_2
s_rmse_2
```

#### Assumptions for Simple Linear Regression Model for placement data-set

```{r}
# Checking for Linearity
ggplot(data = placement_2, aes(x=etest_p, y=mba_p))+geom_point()+geom_smooth(method = "lm")

# Checking for Equal Error Variance
#ggplot(data = s_result_2, aes(x=Predicted, y=s_error_2))+geom_point()
plot(placement_2_model_1, which = 1)

# Checking for Normality of Errors
#qqnorm(s_result_2$s_error_2)
#qqline(s_result_2$s_error_2)
plot(placement_2_model_1, which = 2)
```

#### Multiple Linear Regression Model for placement data-set

```{r}
# Building Model
lm(etest_p~degree_p +mba_p +status, data = placement_2_train) -> placement_2_model_2

# Checking summary of multiple linear regression model (F-Statistics and Adjusted R-Squared)
summary(placement_2_model_2)

# Predicting for test data
predict(placement_2_model_2, newdata = placement_2_test) -> m_pv_2
head(m_pv_2)

# Finding value of error for predictions
cbind(Actual = placement_2_test$etest_p, Predicted = m_pv_2) -> m_result_2
as.data.frame(m_result_2) -> m_result_2

m_result_2$Actual - m_result_2$Predicted -> m_error_2
cbind(m_result_2,m_error_2) -> m_result_2
head(m_result_2)

# Finding root mean squared error (RMSE)
sqrt(mean((m_result_2$m_error_2)^2)) -> m_rmse_2
m_rmse_2
```

#### Assumptions for Multiple Linear Regression Model for placement data-set

```{r}
# Checking for Linearity
ggplot(data = placement_2, aes(x=etest_p, y=degree_p))+geom_point()+geom_smooth(method = "lm")
ggplot(data = placement_2, aes(x=etest_p, y=mba_p))+geom_point()+geom_smooth(method = "lm")
ggplot(data = placement_2, aes(x=etest_p, y=status))+geom_point()+geom_smooth(method = "lm")

# Checking for Equal Error Variance
#ggplot(data = m_result_2, aes(x=Predicted, y=m_error_2))+geom_point()
plot(placement_2_model_2, which = 1)

# Checking for Normality of Errors
#qqnorm(m_result_2$m_error_2)
#qqlint(m_result_2$m_error_2)
plot(placement_2_model_2, which = 2)
```


### (C) Pharmacovigilance data-set

#### Preparing the data

```{r}
# Loading the data-set
pharma_3 <- read.csv('/home/carb0n/Desktop/r/d/Pharmacovigilance_audit_Data.csv')

# Removing LocationID and PatientID columns as they are not needed
pharma_3 <- within(pharma_3, rm(LocationID, PatientID))

# Checking for NA values
colSums(is.na(pharma_3))

# Checking structure of the data-set
str(pharma_3)

# Converting appropriate variables to factor-type
as.factor(pharma_3$Issues) -> pharma_3$Issues
as.factor(pharma_3$DrugID) -> pharma_3$DrugID
as.factor(pharma_3$Gender) -> pharma_3$Gender
```

#### Splitting the data-set into 75:25 ratio based on dependent feature

```{r}
set.seed(1)

sample.split(pharma_3$DrugID, SplitRatio=0.75) -> pharma_3_split_tag
subset(pharma_3, pharma_3_split_tag==T) -> pharma_3_train
subset(pharma_3, pharma_3_split_tag==F) -> pharma_3_test

nrow(pharma_3_train)
nrow(pharma_3_test)
```

#### Simple Linear Regression Model for pharmacovigilance data-set

```{r}
# Building Model
lm(Age~DrugID, data = pharma_3_train) -> pharma_3_model_1

# Checking summary of simple linear regression model (F-Statistics and Adjusted R-Squared)
summary(pharma_3_model_1)

# Predicting for test data
predict(pharma_3_model_1, newdata = pharma_3_test) -> s_pv_3
head(s_pv_3)

# Finding value of error for predictions
cbind(Actual = pharma_3_test$Age, Predicted = s_pv_3) -> s_result_3
as.data.frame(s_result_3) -> s_result_3

s_result_3$Actual - s_result_3$Predicted -> s_error_3
cbind(s_result_3, s_error_3) -> s_result_3
head(s_result_3)

# Finding root mean squared error (RMSE)
sqrt(mean((s_result_3$s_error_3)^2)) -> s_rmse_3
s_rmse_3
```

#### Assumptions for Simple Linear Regression Model for pharmacovigilance data-set

```{r}
# Checking for Linearity
ggplot(data = pharma_3, aes(x=Age, y=DrugID))+geom_point()+geom_smooth(method = "lm")

# Checking for Equal Error Variance
#ggplot(data = s_result_3, aes(x=Predicted, y=s_error_3))+geom_point()
plot(pharma_3_model_1, which = 1)

# Checking for Normality of Errors
#qqnorm(s_result_3$s_error_3)
#qqline(s_result_3$s_error_3)
plot(pharma_3_model_1, which = 2)
```

#### Multiple Linear Regression Model for pharmacovigilance data-set

```{r}
# Building Model 
lm(Age~Issues +DrugID +Gender, data = pharma_3_train) -> pharma_3_model_2

# Checking summary of multiple linear regression model (F-Statistics and Adjusted R-Squared)
summary(pharma_3_model_2)

# Predicting for test data
predict(pharma_3_model_2, newdata = pharma_3_test) -> m_pv_3
head(m_pv_3)

# Finding value of error for predictions
cbind(Actual = pharma_3_test$Age, Predicted = m_pv_3) -> m_result_3
as.data.frame(m_result_3) -> m_result_3

m_result_3$Actual - m_result_3$Predicted -> m_error_3
cbind(m_result_3, m_error_3) -> m_result_3
head(m_result_3)

# Finding root mean squared error (RMSE)
sqrt(mean((m_result_3$m_error_3)^2)) -> m_rmse_3
m_rmse_3
```

#### Assumptions for Multiple Linear Regression Model for pharmacovigilance data-set

```{r}
# Checking for Linearity
ggplot(data = pharma_3, aes(x=Age, y=Issues))+geom_point()+geom_smooth(method = "lm")
ggplot(data = pharma_3, aes(x=Age, y=DrugID))+geom_point()+geom_smooth(method = "lm")
ggplot(data = pharma_3, aes(x=Age, y=Gender))+geom_point()+geom_smooth(method = "lm")

# Checking for Equal Error Variance
#ggplot(data = m_result_3, aes(x=Predicted, y=m_error_3))+geom_point()
plot(pharma_3_model_2, which=1)

# Checking for Normality of Errors
#qqnorm(m_result_3$m_error_3)
#qqline(m_result_3$m_error_3)
plot(pharma_3_model_2, which=2)
```