---
title: "R : Module 1 : Intro To R"
output:
  html_document:
    df_print: paged
header-includes: \usepackage{fvextra} \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      tidy = 'formatR',
                      #tidy.opts = list(blank = FALSE, width.cutoff = 90),
                      prompt = F,
                      size = 'large')
```

### Module 1 : Assignment 1 - Arithmetic Operator

```{r}
# Loading the data-set
churn1 <- read.csv('/home/carb0n/Desktop/r/d/customer_churn.csv')

# a) Increase the monthly charges of 1st record by 5.
churn1$MonthlyCharges[1]+5 -> churn1$MonthlyCharges[1]
churn1$MonthlyCharges[1]

# b) Increase the monthly charges of the 5th record by 22. 
churn1$MonthlyCharges[5]+22 -> churn1$MonthlyCharges[5]
churn1$MonthlyCharges[5]

# c) Reduce the total charges of the 4th record by 40.75.
churn1$TotalCharges[4]-40.75 -> churn1$TotalCharges[4]
churn1$MonthlyCharges[4]

# d) Subtract the total charges from 2nd record by 100. 
churn1$TotalCharges[2]-100 -> churn1$TotalCharges[2]
churn1$MonthlyCharges[2]

# e) Reduce the monthly charges of the 25th record by 4.6 and divide it by 5. 
(churn1$MonthlyCharges[25]-4.6)/5 -> churn1$MonthlyCharges[25]
churn1$MonthlyCharges[25]

# f) Increase the value of the total charge of the 5th record by 5.3 and multiply it by 2. 
(churn1$TotalCharges[5]+5.3)*2 -> churn1$TotalCharges[5]
churn1$MonthlyCharges[5]

# g) Multiply the 12th record of monthly charges by 5. 
churn1$MonthlyCharges[12]*5 -> churn1$MonthlyCharges[12]
churn1$MonthlyCharges[12]
```

### Module 1 : Assignment 2 - Array

```{r}
# Loading the data-set
churn1 <- read.csv('/home/carb0n/Desktop/r/d/customer_churn.csv')

# Loading the data-set
placement1 <- read.csv('/home/carb0n/Desktop/r/d/Placement_Data_Full_Class.csv')

# 1. Create an array as per the given data:

# a. An array named array_total_charges that contains the first 10 values from the total charges column.
array(data=churn1$TotalCharges[1 :10]) -> array_total_charges
array_total_charges

# b. An array named array_monthly_charges that contains first five values from the monthly charges column. 
array(data=churn1$MonthlyCharges[1 :5]) -> array_monthly_charges
array_monthly_charges

# 2. Create an array using placement data set that comprises of: 

# a. Data of first 15 values from salary column. 
array(data=placement1$salary[1 :15]) -> array_salary
array_salary

# b. Data of first five values from mba_p column. 
array(data=placement1$mba_p[1 :5]) -> array_mba_p
array_mba_p

# c. Data of first 10 values from e_test column. 
array(data=placement1$etest_p[1 :10]) -> array_etest_p
array_etest_p
```

### Module 1 : Assignment 3 - C Function

```{r}
# Loading the data-set
pharma1 <- read.csv('/home/carb0n/Desktop/r/d/Pharmacovigilance_audit_Data.csv')

# 1. Extract the following columns from the data set: 

# a. Age and Issues.
df1 <- pharma1[ ,c("Issues","Age")]
head(df1)

# b. Patient ID and Drug ID. 
df2 <- pharma1[ ,c("PatientID", "DrugID")]
head(df2)

# c. Gender, Age, and Issues. 
df3 <- pharma1[ ,c("Gender", "Age", "Issues")]
head(df3)

# 2. Extract the following rows:

# a. Row numbers 1,2,3,4, and 5
pharma1[1 :5,]

# b. Rows from 15 to 75
df4 <- pharma1[15:75,]
head(df4)
tail(df4)

# c. Rows from 100-150
df5 <- pharma1[100:150,]
head(df5)
tail(df5)

# 3. Extract the following rows and columns: 

# a. Columns (1,2,4) and rows(1 to 10)
pharma1[1 :10,c(1,2,4)]

# b. Columns (2,4,5) and rows(50 to 120) 
df6 <- pharma1[50:120,c(2,4,5)]
head(df6)
tail(df6)

# c. Columns (1,3,6) and rows(100 to 200) 
df7 <- pharma1[100:200, c(1,3,6)]
head(df7)
tail(df7)
```

### Module 1 : Assignment 4 - Calculate Mean

```{r}
# Loading the data-set
churn1 <- read.csv('/home/carb0n/Desktop/r/d/customer_churn.csv')

# Loading the data-set
placement1 <- read.csv('/home/carb0n/Desktop/r/d/Placement_Data_Full_Class.csv')

# 1. Calculate the mean of the following columns from customer churn data-set: 

# a. Tenure
mean(churn1$tenure)

# b. Monthly Charges
mean(churn1$MonthlyCharges)

# c. Total Charges
mean(churn1$TotalCharges, na.rm = TRUE)

# 2. Calculate the mean of the following columns from the placement data-set:

# a. Ssc_p
mean(placement1$ssc_p)

# b. Hsc_p
mean(placement1$hsc_p)

# c. Degree_p
mean(placement1$degree_p)

# d. Etest_p
mean(placement1$etest_p)

# e. Mba_p
mean(placement1$mba_p)

# f. Salary 
mean(placement1$salary, na.rm = TRUE)
```

### Module 1 : Assignment 5 - Class Function

```{r}
# Loading the data-set
citytemp1 <- read.csv('/home/carb0n/Desktop/r/d/city_temperature.csv')

# Loading the data-set
placement1 <- read.csv('/home/carb0n/Desktop/r/d/Placement_Data_Full_Class.csv')

# 1. Find the class of the following attributes and covert to character data-type:

# a. Region
class(citytemp1$Region)

# b. Country
class(citytemp1$Country)

# c. City
class(citytemp1$City)

# d. Month
class(citytemp1$Month)
# Converting month to character class
citytemp1$Month <- as.character(citytemp1$Month)

# e. Day
class(citytemp1$Day)
# Converting Day to character class
citytemp1$Day <- as.character(citytemp1$Day)

# f. Year
class(citytemp1$Year)
# Converting Year to character class
citytemp1$Year <- as.character(citytemp1$Year)

# g. AvgTemperature
class(citytemp1$AvgTemperature)
# Converting AvgTemperature to character class
citytemp1$AvgTemperature <- as.character(citytemp1$AvgTemperature)

# 2. After finding the class of the below attributes, convert them to the following class:

# i) Vector to Character

# a. Gender
class(placement1$gender)

# b. ssc_b
class(placement1$ssc_b)

# c. hsc_b
class(placement1$hsc_b)

# d. hsc_s
class(placement1$hsc_s)

# e. status
class(placement1$status)

# f. workex
class(placement1$workex)

# g. specialization 
class(placement1$specialisation)
```

### Module 1 : Assignment 6 - For Loop

```{r}
# Loading the data-set
placement1 <- read.csv('/home/carb0n/Desktop/r/d/Placement_Data_Full_Class.csv')

# Loading the data-set
churn1 <- read.csv('/home/carb0n/Desktop/r/d/customer_churn.csv')

# 1. Using for loop, calculate the number of students from the placement data set according to the following conditions:
# a) Students who are placed.
count_placed=0
for(i in 1 :nrow(placement1)){
  if(placement1$status[i]=="Placed"){
    count_placed=count_placed+1
  }
}
count_placed

# b) Students who are from Science stream.
count_science=0
for(i in 1 :nrow(placement1)){
  if(placement1$degree_t[i]=="Sci&Tech"){
    count_science=count_science+1
  }
}
count_science

# c) Students who are from Commerce stream. 
count_commerce=0
for(i in 1 :nrow(placement1)){
  if(placement1$degree_t[i]=="Comm&Mgmt"){
    count_commerce=count_commerce+1
  }
}
count_commerce



# 2. Using for loop, calculate the number of students from the placement data set who scored more than 80.0 in higher secondary exams (hsc_p).
count_hsc_p80=0
for (i in 1 :nrow(placement1)){
  if(placement1$hsc_p[i]>80.0){
    count_hsc_p80=count_hsc_p80+1
  }
}
count_hsc_p80

# 3. Calculate the number of students who scored more than 75% in MBA (mba_p)and got placed from campus placement drives.
count_mba75_placed=0
for(i in 1 :nrow(placement1)){
  if(placement1$mba_p[i]>75 && placement1$status[i]=="Placed"){
    count_mba75_placed=count_mba75_placed+1
  }
}

count_mba75_placed 

# 4. Calculate the number of senior citizens from customer churn data set who are using internet service as 'DSL'. 
count_senior=0
for(i in 1 :nrow(churn1)){
  if(churn1$SeniorCitizen[i]==1 && churn1$InternetService[i]=="DSL"){
    count_senior=count_senior+1
  }
}
count_senior 
```

### Module 1 : Assignment 7 - If-Else

```{r}
# Loading the data-set
placement1 <- read.csv('/home/carb0n/Desktop/r/d/Placement_Data_Full_Class.csv')

# Loading the data-set
churn1 <- read.csv('/home/carb0n/Desktop/r/d/customer_churn.csv')

# Using the placement data set, perform the following tasks:
# a. Check the value of 95th row of status column.
  # If the value is "Placed", then print "The student is placed from campus placement drives".
  # If it's "Not Placed", then print "The student is not placed from campus placement drives."
  # If it's "NA", then print "The data of the student is not available." 

if(placement1$status[95]=="Placed"){
  print("The student is placed from campus placement drives.")
}else if(placement1$status[95]=="Not Placed"){
  print("The student is not placed from campus placement drives.")
}else(print("The data of the student is not available."))
  

# b. Check the value of 64th row of hsc_s column.
  # If it is "Science", then print "The subject opted by the student is Science."
  # If it is "Commerce", then print "The subject opted by the student is Commerce."
  # If it is "Arts", then print "The subject opted by the student is Arts." 
if(placement1$hsc_s[64]=="Science"){
  print("The subject opted by the student is Science.")
}else if(placement1$hsc_s[64]=="Commerce"){
  print("The subject opted by the student is Commerce.")
}else(print("The subject opted by the student is Arts."))


# c. Check the 28th row of Payment method column:
  # If it is "Credit card (automatic)", then print "The customer is using credit card as their payment method."
  # If it is "Mailed check", then print "The customer is using mailed check for payment."
  # If it is "Electronic check", then print "The customer is using electronic check for payment."
if(churn1$PaymentMethod[28]=="Credit card (automatic)"){
  print("The customer is using credit card as their payment method.")
}else if(churn1$PaymentMethod[28]=="Mailed check"){
  print("The customer is using mailed check for payment.")
}else if(churn1$PaymentMethod[28]=="Electronic check"){
  print("The customer is using electronic check for payment.")
}else{
  print("The customer is using bank transfer as their payment method.")
}
```

### Module 1 : Assignment 8 - If-Else

```{r}
# Loading the data-set
placement1 <- read.csv('/home/carb0n/Desktop/r/d/Placement_Data_Full_Class.csv')

# Loading the data-set
pharma1 <- read.csv('/home/carb0n/Desktop/r/d/Pharmacovigilance_audit_Data.csv')

# Loading the data-set
city1 <- read.csv('/home/carb0n/Desktop/r/d/city_temperature.csv')

# Using the placement data-set, perform the following tasks:
# a. Check the value of 12th row of degree_t column, if the value is "Comm&Mgmt", and then print "The student is of the stream Commerce and Management".
if(placement1$degree_t[12]=="Comm&Mgmt"){
  print("The student is of the stream Commerce and Management")
}

# b. Check the value of 23rd row of specialization column, if the value is "Mkt&HR", then print "The specialization of student is Marketing and Human Resource"
if(placement1$specialisation[23]=="Mkt&HR"){
  print("The specialization of student is Marketing and Human Resource")
}

# c. Check the 39th row of status column, if the value is "Placed", then print "The student is placed from campus."
if(placement1$status[39]=="Placed"){
  print("The student is placed from campus.")  
}


# Using the pharmacovigilance data-set, perform the following tasks:
# a. Check the value of 37th row of Issues column, if the value is "Medication history documenting error", then print "The patient's medical record is missing"
if(pharma1$Issues[37]=="Medication history documenting error"){
  print("The patient's medical record is missing")
}

# b. Check the value of 180th row of Issues column, if the value is "Unclear Route", then print "The patient is having an unclear route of transmission"
if(pharma1$Issues[180]=="unclear route"){
  print("The patient is having an unclear route of transmission")
}

# c. . Check the 220th row of Location ID column, if the value is "Location3", then print "The location ID of the patient is Location3."
if(pharma1$LocationID[220]=="Location3"){
  print("The location ID of the patient is Location3.")  
}

# Using the city-temperature data-set, perform the following tasks:
# a. Check the value of 11th row of Region column, if the value is "Africa", then print "The region for which we are calculating average temperature is Africa"
if(city1$Region[11]=="Africa"){
  print("The region for which we are calculating average temperature is Africa")
}

# b. Check the value of row number 79961 of city column, if the value is "Conakry", then print "The city for which we are calculating average temperature is Conakry".
if(city1$City[79961]=="Conakry"){
  print("The city for which we are calculating average temperature is Conakry")
}

# c. Check the value of 70th row of year column, if the value is "1995", then prints "We are calculating average temperature for the year 1995".
if(city1$Year[70]=="1995"){
  print("We are calculating average temperature for the year 1995")  
}
```

### Module 1 : Assignment 9 - Switch Case

```{r}
# Loading the data-set
placement1 <- read.csv('/home/carb0n/Desktop/r/d/Placement_Data_Full_Class.csv')

# Loading the data-set
churn1 <- read.csv('/home/carb0n/Desktop/r/d/customer_churn.csv')

# 1. Using switch case, check 67th row of hsc_s column from placement data:
  # If it is "Science", then add 5 extra marks in degree_p score.
  # If it is "Commerce", then add 3 extra marks in degree_p score. 
switch(as.character(placement1$hsc_s[67]),
       "Science"=placement1$degree_p[67]+5,
       "Commerce"=placement1$degree_p[67]+3) -> placement1$degree_p[67]
placement1$degree_p[67]

# 2. Check 74th row of ssc_b column in the placement data set:
  #  If it is "Others", then add 3 as grace marks in hsc_p for the 74th row.
  #  If it is "Central", then add 5 as grace marks in hsc_p for the 74th row.S
switch(as.character(placement1$ssc_b[74]),
       "Others"=placement1$hsc_p[74]+3,
       "Central"=placement1$hsc_p[74]+5) -> placement1$hsc_p[74]
placement1$hsc_p[74]

# 3. Using switch case, check 4th row of contract column from customer churn data:
#  If it is "One Year", then give a 20% discount in total charges.
#  If it is "Month-to-month", then give a discount of 5% in total charges. 
switch(as.character(churn1$Contract[4]),
       "One year"=churn1$TotalCharges[4]-(churn1$TotalCharges[4] * 0.20),
       "Month-to-month"=churn1$TotalCharges[4]-(churn1$TotalCharges[4] * 0.05)) -> churn1$TotalCharges[4]
churn1$TotalCharges[4]

# 4. Check 14th row of Internet Service column in the customer churn data set:
#  If the customer is using "Fiber optic", then give a discount of 15% in total charges.
#  If it is "DSL", then give a discount of 10% in total charges. 
switch(as.character(churn1$InternetService[14]),
       "Fiber optic"=churn1$TotalCharges[14]-(churn1$TotalCharges[14] * 0.15),
       "DSL"=churn1$TotalCharges[14]-(churn1$TotalCharges[14] * 0.10)) -> churn1$TotalCharges[14]
churn1$TotalCharges[14]
```

### Module 1 : Assignment 10 - While Loop

```{r}
# Loading the data-set
placement1 <- read.csv('/home/carb0n/Desktop/r/d/Placement_Data_Full_Class.csv')

# Loading the data-set
churn1 <- read.csv('/home/carb0n/Desktop/r/d/customer_churn.csv')

# 1. Calculate the number higher secondary students who are from "Central" board.
count_ssc_central=0
i=1
while(i<=nrow(placement1)){
  if(placement1$ssc_b[i]=="Central"){
    count_ssc_central=count_ssc_central+1  
  }
i=i+1
}
count_ssc_central

# 2. Calculate the number of students who are from Science stream and secured percentage greater than 75. 
count_hsc_science=0
i=1
while(i<=nrow(placement1)){
  if(placement1$hsc_s[i]=="Science" && placement1$hsc_p[i] > 75){
    count_hsc_science=count_hsc_science+1  
  }
  i=i+1
}
count_hsc_science

# 3. Calculate the number of customers from customer churn data set as per following condition:
#  (i) The customer should be a senior citizen who is having a contract of 'One year' for 'Fiber optic' there as internet service.
count=0
i=1
while(i<=nrow(churn1)){
  if(churn1$SeniorCitizen[i]==1 && churn1$Contract[i]=="One year" && churn1$InternetService[i]=="Fiber optic"){
    count=count+1  
  }
  i=i+1
}
count

#(ii) The customer is not a senior citizen who is having a contract of 'Two year' and total charges is less than 7000.

#Checking for NA values
sum(is.na(churn1))

#Removing NA Values
churn1 <- na.omit(churn1)

count=0
i=1
while(i<=nrow(churn1)){
  if(churn1$SeniorCitizen[i]==0 && churn1$Contract[i]=="Two year" && churn1$TotalCharges[i] < 7000){
    count=count+1  
  }
  i=i+1
}
count

#(iii) The customers whose monthly charges are greater than 6500. 
count=0
i=1
while(i<=nrow(churn1)){
  if(churn1$MonthlyCharges[i] > 6500){
    count=count+1  
  }
  i=i+1
}
count
```

### Module 1 : Assignment 11 - Subset Function

```{r}
# Loading the data-set
pharma1 <- read.csv('/home/carb0n/Desktop/r/d/Pharmacovigilance_audit_Data.csv')

# 1. Extract a subset of pharmacovigilance data of all the female patients whose age is less than 25 with issues of unclear dose. 
df1 <- subset(pharma1, Gender=="F" & Age <=25 & Issues=="unclear dose")
head(df1)

# 2. Extract the records of all the male patients whose age is less than 25 and medication data is not available. 
df2 <- subset(pharma1, Gender=="M" & Age <=25 & Issues=="Medication history documenting error")
head(df2)

# 3. Fetch the records of children whose age equals to 8 whose documentation of medication history is not available. 
df3 <- subset(pharma1, Age==8 & Issues=="Medication history documenting error")
head(df3)

# 4. Extract all the records of female children of age>5 having an unclear dose.
df4 <- subset(pharma1, Gender=="F" & Age>5 & Issues=="unclear dose")
head(df4)

# 5. Extract the records of male patients of age equals to 35 with an unclear route of transmission. 
df5 <- subset(pharma1, Gender=="M" & Age==35 & Issues=="unclear dose")
head(df5)
```

### Module 1 : Assignment 12 - Calculating Number of Rows

```{r}
# Loading the data-set
placement1 <- read.csv('/home/carb0n/Desktop/r/d/Placement_Data_Full_Class.csv')

# 1. Check how many rows and columns are there in the data-set?
nrow(placement1)
ncol(placement1)

# 2. Extract the data of people who are having the branch as Sci&Tech and who are employed.  
# Then, check how many such people are there who are employed by finding the number of rows that contains the data of employed people.

df1 <- subset(placement1, degree_t=="Sci&Tech" & status=="Placed")
nrow(df1)

# 3. Find the number of rows of female students who are from Arts or Commerce stream. 
df2 <- subset(placement1, gender=="F" & (hsc_s=="Arts" | hsc_s=="Commerce"))
nrow(df2)

# 4. Fetch the number records of all the students as per the following condition and find the number of such students:

# a. Students who are placed
df3 <- subset(placement1, status=="Placed")
nrow(df3)

# b. Students who are earning more than 300000
df4 <- subset(placement1, salary > 300000)
nrow(df4)

# 5. Find the number of columns having data type as following:
# a. Integer
# b. Factor
table(sapply(placement1,class))

#OR
str(placement1)
```

### Module 1 : Assignment 13 - Logical Operator

```{r}
# Loading the data-set
churn1 <- read.csv('/home/carb0n/Desktop/r/d/customer_churn.csv')

# 1. Extract the data of customers as per the following conditions:

# a. People who are senior citizen of gender female.
df1 <- subset(churn1, SeniorCitizen==1 & gender=="Female")
head(df1)

# b. Female customers who are not having partner.
df2 <- subset(churn1, gender=="Female" & Partner=="No")
head(df2)

# c. Customers using 'DSL' as their internet service with no online security.
df3 <- subset(churn1, InternetService=="DSL" & OnlineSecurity=="No")
head(df3)

# d. Customers using 'Fiber optics' as internet service with no online backup.
df4 <- subset(churn1, InternetService=="Fiber optic" & OnlineBackup=="No")
head(df4)

# e. Senior citizens using payment method as electronic check. 
df5 <- subset(churn1, SeniorCitizen==1 & PaymentMethod=="Electronic check")
head(df5)  
  

# 2. Extract the data of customers who are:

# a. Using payment method as mailed check and having a contract of one year
df6 <- subset(churn1, PaymentMethod=="Mailed check" & Contract=="One year")
head(df6)

# b. Having services such as online security, online backup, device protection, and tech support enabled
df7 <- subset(churn1, OnlineSecurity=="Yes" & OnlineBackup=="Yes" & DeviceProtection=="Yes" & TechSupport=="Yes")
head(df7)

# c. Not senior citizens having tenure of 72 months.  
df8 <- subset(churn1, SeniorCitizen==0 & tenure==72)
head(df8)  


# 3. Check the customers data where the customer's are:

# a. Either using bank transfer or credit card as their payment method.
df9 <- subset(churn1, PaymentMethod=="Bank transfer (automatic)" | PaymentMethod=="Credit card (automatic)")
head(df9)

# b. Having phone service or multiple lines enabled.
df10 <- subset(churn1, PhoneService=="Yes")
head(df10)

# c. Either having payment method as mailed check or electronic check.
df11 <- subset(churn1, PaymentMethod=="Mailed check" | PaymentMethod=="Electronic check")
head(df11)

# d. Either using fiber optic or having no internet service.   
df12 <- subset(churn1, InternetService=="Fiber optic" | InternetService=="No")
head(df12) 
```

### Module 1 : Assignment 14 - Relational Operator

```{r}
# Loading the data-set
placement1 <- read.csv('/home/carb0n/Desktop/r/d/Placement_Data_Full_Class.csv')

# 1. Extract the data of the students who are:

# a. Having SSC score greater than or equal to 65.
df1 <- subset(placement1, ssc_p >= 65)
head(df1)

# b. Having HSC score greater than 80.
df2 <- subset(placement1, hsc_p > 80)
head(df2)

# c. Earning more than 200000. 
df3 <- subset(placement1, salary > 200000)
head(df3)


# 2. Extract the data of the students who are:

# a. Placed and earning salary less than 250000.
df4 <- subset(placement1, status=="Placed" & salary < 250000)
head(df4)

# b. From commerce and management having earning salary less than 600000.
df5 <- subset(placement1, degree_t=="Comm&Mgmt" & salary < 600000)
head(df5)

# c. From Science stream with etest score greater than 90.     
df6 <- subset(placement1, hsc_s=="Science" & etest_p > 90)
head(df6)


# 3. Extract the data of the students who are:

# a. From Central board studying Arts.
df7 <- subset(placement1, ssc_b=="Central" & hsc_s=="Arts")
head(df7)

# b. From Sci&Tech with e_test score equals to 75.
df8 <- subset(placement1, degree_t=="Sci&Tech" & etest_p==75)
head(df8)  

# c. Placed with a salary of 450000.   
df9 <- subset(placement1, status=="Placed" & salary==450000)
head(df9)
```

### Module 1 : Assignment 15 - table()

```{r}
# Loading the data-set
churn1 <- read.csv('/home/carb0n/Desktop/r/d/customer_churn.csv')

# 1. Check the categories of various columns given below using table function :

# a. Gender
table(churn1$gender)
  
# b. Senior Citizen
table(churn1$SeniorCitizen)
  
# c. Partner
table(churn1$Partner)
  
# e. Dependents
table(churn1$Dependents)
  
# f. Phone Service
table(churn1$PhoneService)
  
# g. Multiple Lines
table(churn1$MultipleLines)
  
# h. Internet Service
table(churn1$InternetService)
  
# i. Online Security
table(churn1$OnlineSecurity)
  
# j. Online Backup
table(churn1$OnlineBackup)
  
# k. Contract
table(churn1$Contract)
  
# l. Payment Method 
table(churn1$PaymentMethod)

# And extract the data of male customers who are senior citizens with partners, internet service as DSL and phone services enabled     
df1 <- subset(churn1, gender=="Male" & SeniorCitizen==1 & Partner=="Yes" & InternetService=="DSL" & PhoneService=="Yes")
head(df1)
```

### Module 1 : Assignment 16 - head()

```{r}
# Loading the data-set
placement1 <- read.csv('/home/carb0n/Desktop/r/d/Placement_Data_Full_Class.csv')

# 1. Check the first five values of the following columns using the head function:
# a. Sl_no
# b. Gender
# c. Ssc_p
# d. Ssc_b
# e. Hsc_p
# f. Degree_p
# g. Degree_t
# h. Etest_p
# i. Specialization
# j. Mba_p
# k. Salary
head(placement1)

# 2. Add 5000 to the first six data of the salary column. 
head(placement1$salary,6)+5000

# 3. Add 7.5 to the first 10 values of e_test column. 
head(placement1$etest_p,10)+7.5

# 4. Reduce the mba_p score by 5 for the first three values
head(placement1$mba_p,3)-5
```

### Module 1 : Assignment 17 - tail()

```{r}
# Loading the data-set
placement1 <- read.csv('/home/carb0n/Desktop/r/d/Placement_Data_Full_Class.csv')

# 1. Extract the last few values of the following columns from the placement data-set:
# a. Ssc_p
# b. Hsc_p
# c. Degree_p
# d. Etest_p
# e. Mba_p
# f. Salary
tail(placement1)

# 2. Check and add 12000 to the last 8 values of the salary column.
tail(placement1$salary,8)+12000

# 3. Check and increase the mba_p score by 10 for the last few values. 
tail(placement1$mba_p)
tail(placement1$mba_p)+10
```

### Module 1 : Assignment 18 - sample()

```{r}
# Loading the data-set
pharma1 <- read.csv('/home/carb0n/Desktop/r/d/Pharmacovigilance_audit_Data.csv')

# 1. Extract some random set of values using sample function.

# a. Select five random values from LocationID.
sample(pharma1$LocationID, size=5)

# b. Extract 25 random values from age column.
sample(pharma1$Age, size=25)

# c. Extract 15 random from PatientID.
sample(pharma1$PatientID, size=15)

# d. Extract 10 random values from Issues column. 
sample(pharma1$Issues, size=10)
```

### Module 1 : Assignment 19 - str()

```{r}
# Loading the data-set
churn1 <- read.csv('/home/carb0n/Desktop/r/d/customer_churn.csv')

# Loading the data-set
placement1 <- read.csv('/home/carb0n/Desktop/r/d/Placement_Data_Full_Class.csv')

# Loading the data-set
pharma1 <- read.csv('/home/carb0n/Desktop/r/d/Pharmacovigilance_audit_Data.csv')

# Loading the data-set
city1 <- read.csv('/home/carb0n/Desktop/r/d/city_temperature.csv')

# 1. Find the structure of customer churn data-set using str() function and perform the following operations:
str(churn1)
  
# a. Convert the data type of internet service column from vector to character.
as.character(churn1$InternetService) -> churn1$InternetService
  
# b. Convert the data type of partner column from vector to character.
as.character(churn1$Partner) -> churn1$Partner
  
# c. Convert the data type of total charges column from num to int.
as.integer(churn1$TotalCharges) -> churn1$TotalCharges


# 2. Find the structure of the placement data.
str(placement1)
  
# 3. Find the structure of the pharmacovigilance data.
str(pharma1)  

# 4. Find the structure of city temperature data-set. 
str(city1)  
```

### Module 1 : Assignment 20 - Matrix

```{r}
# 1. Create a matrix named 'm_even' with the 16 continuous even values. The matrix should have 4 rows & 4 columns. Also, arrange the elements by row.
m_even <- matrix(data=seq(2, by=2, len=16), nrow=4, ncol=4, byrow = TRUE)
m_even

# 2. Create a matrix name 'm_odd' having 4 rows and 4 columns with 16 consecutive odd values arranged by row.
m_odd <- matrix(data=seq(1, by=2, len=16), nrow=4, ncol=4, byrow = TRUE)
m_odd

# 3. Create a matrix named m_8 as per the following conditions:
# a. The dimension of the matrix should be 8x8.
# b. It should have consecutive values starting from 1.
# c. It should be arranged by row. 
m_8 <- matrix(data=seq(1, by=1, len=64), nrow=8, ncol=8, byrow=TRUE)
m_8
```

### Module 1 : Assignment 21 - Lists

```{r}
# 1. Create a list named internet_service using the following condition:

# a. A character vector that consists of name of the internet services as "DSL", "Fiber Optic", "Cable Broadband", and "Wireless".
vec_1 <- c("DSL", "Fiber Optic", "Cable Broadband", "Wireless")  
class(vec_1)

# b. A numeric vector that consists of values from 30 to 40
vec_2 <- c(seq(30,by=1,40))
class(vec_2)

# c. A logical vector having 6 values as: True, False, True, True, True, and False
vec_3 <- c(TRUE, FALSE, T, T, T, F)
class(vec_3)

# List named internet_service including above vectors
internet_service <- list(vec_1, vec_2, vec_3)
internet_service

# 2. Create a list named payment_method using the following condition:

# a. A character vector comprising of payment methods as: 'Electronic Check', 'Mailed Check', 'Credit Card', and 'Online Transfer'.
v1 <- c("Electronic Check", "Mailed Check", "Credit Card", "Online Transfer")
v1

# b. A numeric vector that consists of values from 85-95
v2 <- c(seq(85,by=1,95))
v2

# c. A logical vector having 6 values as: True, True, False, True, False, and True. 
v3 <- c(T,T,F,T,F,T)
v3

# List named payment_method including above vectors
payment_method <- list(v1, v2, v3)
payment_method

# 3. Create a list named contract_list using the following condition:

# a. A character vector comprising of contract type as: "Month to month", "One Year", "Two Year", and "Three Year".
vect_1 <- c("Month to month", "One Year", "Two Year", "Three Year")
vect_1

# b. A numeric vector comprising of five values as: '1', '12', '17', '32', and '72'.
vect_2 <- c(1,12,17,32,72)
vect_2

# c. A logical vector comprising of four values: True, False, True, and True.
vect_3 <- c(T,F,T,F) 
vect_3

# List named internet_service including above vectors
contract_list <- list(vect_1,vect_2,vect_3)
contract_list

# i. Now, extract the 2nd value from the first element of the list
contract_list[[1]][[2]]
  
# ii. Extract the 3rd value from the 2nd element of the list
contract_list[[2]][[3]]

# iii. Extract the 4th value from the 1st element of the lists 
contract_list[[1]][[4]]
```

### Module 1 : Assignment 22 - Vectors

```{r}
# 1. Create the character vector:
# a. Name it as v_specialization
# b. Put values as: "Marketing", "Finance", "TechSupport", and "Development"
v_specialization <- c("Marketing", "Finance", "TechSupport", "Development")
v_specialization

# 2. Create the numeric vector:
# a. Name it as v_num
# b. Put values as: 2, 4, 16, 64, 256, and 1024.
v_num <- c(2,4,16,64,256,1024)
v_num

# 3. Create the logical vector:
# a. Name it as v_logic
# b. Put values as: True, False, True, True, and True. 
v_logic <- c(T,F,T,T,T)
v_logic
```

### Module 1 : Assignment 23 - range()

```{r}
# Loading the data-set
churn1 <- read.csv('/home/carb0n/Desktop/r/d/customer_churn.csv')

# Loading the data-set
placement1 <- read.csv('/home/carb0n/Desktop/r/d/Placement_Data_Full_Class.csv')

# Loading the data-set
pharma1 <- read.csv('/home/carb0n/Desktop/r/d/Pharmacovigilance_audit_Data.csv')

# Loading the data-set
city1 <- read.csv('/home/carb0n/Desktop/r/d/city_temperature.csv')


# 1.Calculate the range of following columns of customer churn dataset:

# a.Tenure
range(churn1$tenure)

# b.Monthly Charges
range(churn1$MonthlyCharges)

# c.Total Charges
range(churn1$TotalCharges, na.rm=TRUE)
  

# 2.Calculate the range of following columns of city temperature dataset:

# a. Day
range(city1$Day)
  
# b.Year
range(city1$Year)
  
# c.Average Temperature
range(city1$AvgTemperature)
  

# 3.Calculate the range of following columns of pharmacovigilance dataset:

# a.Age
range(pharma1$Age)

# b.Patient ID
range(pharma1$PatientID)


# 4.Calculate the range of following columns of placement dataset:

# a.Sl_no
range(placement1$sl_no)
  
# b.Ssc_p
range(placement1$ssc_p)
  
# c.Hsc_p
range(placement1$hsc_p)
  
# d.Degree_p
range(placement1$degree_p)
  
# e.Etest_p
range(placement1$etest_p)
  
# f.Mba_p
range(placement1$mba_p)
  
# g.Salary 
range(placement1$salary, na.rm=TRUE)
```

### Module 1 : Case Study 1 - Data Exploration

```{r}

# Loading the data-set
customer_churn <- read.csv('/home/carb0n/Desktop/r/d/customer_churn.csv')

# 1. Extract these individual columns with the '$' symbol:

# a. Extract 'InternetService' column and store it in'customer_internet_service'
customer_churn$InternetService -> customer_internet_service 
head(customer_internet_service)

# b. Extract 'PaperlessBilling' column and store it in'customer_Paperless_Billing'
customer_churn$PaperlessBilling -> customer_Paperless_Billing 
head(customer_Paperless_Billing)

# c. Extract 'StreamingTV' column and store it in'customer_Streaming_TV'
customer_churn$StreamingTV -> customer_Streaming_TV 
head(customer_Streaming_TV)


# 2. Extract the 3rd, 6th and 9th columns from the 'customer_churn' data.frame & store it in 'customer_random_columns'
customer_churn[,c(3,6,9)] -> customer_random_columns
head(customer_random_columns)


# 3. Extract all the columns from column number-10 to column-number 20 and store the result in 'customer_10_20'
customer_churn[,(10:20)] -> customer_10_20
head(customer_10_20)


# 4. Extract only these row numbers: 65, 765, 3726 & 7000 and store the result in 'customer_random_rows'
customer_churn[c(65,765,3726,7000),] -> customer_random_rows
customer_random_rows


# 5. Extract all the rows starting from row number-655 to row number-6550 & store the result in'customer_continuous_rows'
customer_churn[(655:6550),] -> customer_continuous_rows
head(customer_continuous_rows)
tail(customer_continuous_rows)


# 6. Extract row numbers- 10, 100 & 1000 & column numbers- 5, 10, 15 & store the result in 'customer_random_data'
customer_churn[c(10,100,1000),c(5,10,15)] -> customer_random_data
customer_random_data
```

### Module 1 : Case Study 2 - Operators

```{r}
# Loading the data-set
customer_churn <- read.csv('/home/carb0n/Desktop/r/d/customer_churn.csv')

# 1. Working with arithmetic operators:

# a. Add 5 to the fifth record of 'MonthlyCharges' column
customer_churn$MonthlyCharges[5]+5 -> customer_churn$MonthlyCharges[5]
customer_churn$MonthlyCharges[5]

# b. Subtract 9.65 from the sixth record of 'MonthlyCharges' column
customer_churn$MonthlyCharges[6]-9.65 -> customer_churn$MonthlyCharges[6]
customer_churn$MonthlyCharges[6]

# c. Multiply the 1st record of 'MonthlyCharges' column with 3
customer_churn$MonthlyCharges[1]*3 -> customer_churn$MonthlyCharges[1] 
customer_churn$MonthlyCharges[1] 

# d. Divide the 37th record of 'MonthlyCharges' column with 3
customer_churn$MonthlyCharges[37]/3 -> customer_churn$MonthlyCharges[37]
customer_churn$MonthlyCharges[37]


# 2. Working with the relational operators:
# a. Check if the value of 'tenure' in the 1st row is greater than the value of 'tenure' in the 10th row
customer_churn$tenure[1] > customer_churn$tenure[10] 
 
# b. Check if the value of 'tenure' in the 3rd row is equal to the value of 'tenure' in the 5th row
customer_churn$tenure[3] == customer_churn$tenure[5]


# 3. Working with logical operators:
# a. Get the count of those customers who have subscribed to both "TechSupport" &"StreamingTV"
nrow(subset(customer_churn, customer_churn$TechSupport=="Yes" & customer_churn$StreamingTV=="Yes")) 

# b. Extract those customers whose 'InternetService' is either 'DSL' or 'Fiber optic' & store the result in 'Internet_dsl_fiber
Internet_dsl_fiber <- subset(customer_churn, InternetService=="DSL" | InternetService=="Fiber optic") 
head(Internet_dsl_fiber) 
```

### Module 1 : Case Study 3 - In-Built Functions

```{r}
# Loading the data-set
customer_churn <- read.csv('/home/carb0n/Desktop/r/d/customer_churn.csv')


# 1. Do these operations with the head() function:
# a. Get the first 4 records from 'PhoneService' column
head(customer_churn$PhoneService,4)  

# b. Get the first 8 records from 'Contract' column
head(customer_churn$Contract,8)


# 2. Do these operations with the tail() function:

# a. Get the last record of 'TotalCharges' column
tail(customer_churn$TotalCharges,1)
  
# b. Get the last 5 records of 'tenure' column
tail(customer_churn$tenure,5)

# 3. Find the average, minimum, maximum & range from the 'tenure' column
summary(customer_churn$tenure)
range(customer_churn$tenure)
# OR
mean(customer_churn$tenure)
min(customer_churn$tenure)
max(customer_churn$tenure)

# 4. Get 10 random values from the 'TotalCharges' column using sample()
sample(customer_churn$TotalCharges, size=10)

# 5. Find the count of different levels in 'PaymentMethod' & 'Contract' columns using table() 
table(customer_churn$PaymentMethod)
table(customer_churn$Contract)
```

### Module 1 : Case Study 4 - Data Structures

```{r}
# Loading the data-set
customer_churn <- read.csv('/home/carb0n/Desktop/r/d/customer_churn.csv')

# 1. Create these vectors:

# a. A character vector named 'fruits' with these values: 'Apple', 'Guava', 'Banana', 'Mango'
fruits <- c("Apple", "Guava", "Banana", "Mango")
fruits

# b. A numeric vector named 'hundred' comprising of the first 100 natural numbers
hundred <- c(seq(1, by=1, 100))
hundred

# c. A logical vector named 'logic_game' with these values: 'TRUE','TRUE','FALSE','FALSE'
logic_game <- c(T,T,F,F)
logic_game

# 2. Create a list named 'jumbo' which comprises of:

# a. A character vector comprising of alphabets from A to D
item_1 <- (toupper(letters[1 :4]))
item_1
  
# b. A numeric vector comprising of numbers from 55 to 60
item_2 <- c(seq(55,by=1,60))
item_2
  
# c. A logical vector comprising of just these two values: True, False
item_3 <- c(T,F)
item_3  

# List named 'jumbo' which comprises of above vectors
jumbo <- list(item_1,item_2,item_3) 
jumbo

# i. Now, access the third value from the first element of the list
jumbo[[1]][[3]]  

# ii. Access the 2nd value from the 2nd element of the list
jumbo[[2]][[2]]

# iii. Access the 1st value from the 3rd element of the list
jumbo[[3]][[1]]


# 3. Create a matrix named 'four_trouble', with the numbers 1 to 16.  The matrix should have 4 rows & 4 columns
four_trouble <- matrix(data=seq(1,by=1,16), nrow=4, ncol=4)  
four_trouble  

# a. Arrange the elements by row
four_trouble <- matrix(data=seq(1,by=1,16), nrow=4, ncol=4, byrow = TRUE)
four_trouble

# 4. Create an array named 'sky_maze' with the numbers 1 to 32.  The array should comprise of two 4*4 matrices 
sky_maze <- array(data=seq(1,by=1,32), dim=c(4,4,2))
sky_maze 
```

### Module 1 : Case Study 5 - Flow Control

```{r}
# Loading the data-set
churn1 <- read.csv('/home/carb0n/Desktop/r/d/customer_churn.csv')


# 1. Check if the value in the 6th cell of 'PaymentMethod' column is 'Electronic check'. If yes, print "Yes, the payment method is Electronic check"
if(churn1$PaymentMethod[6]=="Electronic check"){
  print("Yes, the payment method is Electronic check")
}

# 2. Check the value present in 12th cell of 'Contract' column.
# If it's 'month-to-month', print 'The contract is on a month to month basis'
# If it's 'One year', print 'The contract is on a yearly basis'
# If it's 'Two year', print 'The contract is on a two-year basis'
if(churn1$Contract[12]=="Month-to-month"){
  print("The contract is on a month to month basis")
}else if(churn1$Contract[12]=="One year"){
  print("The contract is on a yearly basis")
}else if(churn1$Contract[12]=="Two year"){
  print("The contract is on a two-year basis")
}

# 3. Use switch to check the gender in 6th cell of 'gender' column.
# If it's 'Male', give a discount of 20% in 'MonthlyCharges'
# If it's 'Female', give a discount of 50% in 'MonthlyCharges'
switch(as.character(churn1$gender[6]),
       "Male"=churn1$MonthlyCharges[6]-(churn1$MonthlyCharges[6]*0.2),
       "Female"=churn1$MonthlyCharges[6]-(churn1$MonthlyCharges[6]*0.5)) -> churn1$MonthlyCharges[6]
churn1$MonthlyCharges[6]

# 4. Use for loop to get the count of customers whose 'InternetService' is'DSL'
count=0
for(i in 1 :nrow(churn1)){
  if(churn1$InternetService[i]=="DSL"){
    count=count+1
  }
}
count

# 5. Use while to find the number of customers whose tenure is exactly '2' months
count=0
i=1
while(i<=nrow(churn1)){
  if(churn1$tenure[i]==2){
    count=count+1
  }
  i=i+1
}
count
```