---
title: "R : Module 9 : Unsupervised Learning"
output:
  html_document:
    df_print: paged
header-includes: \usepackage{fvextra} \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      tidy = 'formatR',
                      tidy.opts = list(blank = FALSE, width.cutoff = 90),
                      prompt = F,
                      size = 'large')
```

## Module 9 : Assignment 1-4

### Loading required packages

```{r}
library('dplyr')
library('caTools')
```

#### Task-1

```{r}
#a. Build the kmeans algorithm on top of 'customer_features'. For the model, the number of clusters should be 3.
#b. Calculate the clustering vector values for the monthly charges column form the customer_features.
#c. Bind the monthly charges column to the clustering vector and store that data in month_group.
#d. Convert the month_group matrix into a data frame.
#e. Separate all the 3 clusters with their values. 
```

#### Preparing the data

```{r}
# Loading the data-set
churn_4 <- read.csv('/home/carb0n/Desktop/r/d/customer_churn.csv')

# Checking structure of data-set
str(churn_4)

# Checking for NA values
colSums(is.na(churn_4))

# Removing "CustomerID" column as it is not needed
churn_4 <- within(churn_4, rm(customerID))

# Plotting box-plot for TotalCharges column to check for outliers
boxplot(churn_4$TotalCharges)

# Replacing NA values in TotalCharges column with mean value for TotalCharges column
churn_4$TotalCharges[is.na(churn_4$TotalCharges)] <- mean(churn_4$TotalCharges, na.rm = TRUE)

# Converting appropriate variables to factor-type 
as.factor(churn_4$gender) -> churn_4$gender
as.factor(churn_4$SeniorCitizen) -> churn_4$SeniorCitizen
as.factor(churn_4$Partner) -> churn_4$Partner
as.factor(churn_4$Dependents) -> churn_4$Dependents
as.factor(churn_4$PhoneService) -> churn_4$PhoneService
as.factor(churn_4$MultipleLines) -> churn_4$MultipleLines
as.factor(churn_4$InternetService) -> churn_4$InternetService
as.factor(churn_4$OnlineSecurity) -> churn_4$OnlineSecurity
as.factor(churn_4$OnlineBackup) -> churn_4$OnlineBackup
as.factor(churn_4$DeviceProtection) -> churn_4$DeviceProtection
as.factor(churn_4$TechSupport) -> churn_4$TechSupport
as.factor(churn_4$StreamingTV) -> churn_4$StreamingTV
as.factor(churn_4$StreamingMovies) -> churn_4$StreamingMovies
as.factor(churn_4$Contract) -> churn_4$Contract
as.factor(churn_4$PaperlessBilling) -> churn_4$PaperlessBilling
as.factor(churn_4$PaymentMethod) -> churn_4$PaymentMethod
as.factor(churn_4$Churn) -> churn_4$Churn
```

#### K-Means Model  

```{r}
# Selecting specific columns to create data-frame
churn_4 %>% select("tenure", "MonthlyCharges", "TotalCharges") -> customer_features

# Building kmeans algorithm on top of data-frame
kmeans(customer_features$MonthlyCharges, 3) -> k_month

# Checking clustering vector values
head(k_month$cluster)

# Binding clustering vector values to column and converting to data-frame
cbind(Month = customer_features$MonthlyCharges, Cluster = k_month$cluster) -> month_group
as.data.frame(month_group) -> month_group
head(month_group)
```

#### Separating Clustering Vector Values

```{r}
# Separating clustering vector values for each cluster
month_group %>% filter(Cluster==1) -> month_group_1
month_group %>% filter(Cluster==2) -> month_group_2
month_group %>% filter(Cluster==3) -> month_group_3

# Checking separated cluster values for each cluster
head(month_group_1)
head(month_group_2)
head(month_group_3)
```

## Module 9: Case Study 1 - Recommender System 

### Loading required packages

```{r}
library('dplyr')
library('reshape2')
library('recommenderlab')
```

#### Preparing the data

```{r}
# Loading the data-set
ratings <- read.csv('/home/carb0n/Desktop/r/d/movielense/ratings.csv')

# Loading the data-set
movies <- read.csv('/home/carb0n/Desktop/r/d/movielense/movies.csv', stringsAsFactors = FALSE)
```

#### Separating all the movies rated less often

```{r}
# Selecting movies rated less often (only include top 30% rated movies)
df_movie_summary = ratings %>% group_by(movieId) %>% summarise(count = n(),mean(rating))
movie_benchmark = round(quantile(df_movie_summary$count,0.70, names = FALSE),0)
drop_movie_list = df_movie_summary[df_movie_summary$count < movie_benchmark,1]
drop_movie_list = as.data.frame(drop_movie_list)
head(drop_movie_list)
```

#### Separating all the users who rate less often

```{r}
# Creating data-frame of all the inactive users (users who rate less often)
df_movie_summary1 = ratings %>% group_by(userId) %>% summarise(count = n(),mean(userId))
cust_benchmark = round(quantile(df_movie_summary1$count,0.70, names = FALSE),0)
drop_cust_list = df_movie_summary1[df_movie_summary1$count < cust_benchmark,1]
drop_cust_list = as.data.frame(drop_cust_list)
head(drop_cust_list)
```

#### Removing all the movies rated less often and the users who rate less often

```{r}
ratings = anti_join(ratings, drop_movie_list, by = "movieId")
ratings = anti_join(ratings, drop_cust_list, by = "userId")
```

#### Reshaping the data

```{r}
# Creating ratings matrix for movie-ratings with Rows = userId, Columns = movieId
ratingmat <- dcast(ratings, userId~movieId, value.var = "rating", na.rm=FALSE)

# Removing UserID column as it is not needed
ratingmat <- as.matrix(ratingmat[,-1]) #remove userIds

# Converting rating matrix into a recommenderlab sparse matrix
ratingmat <- as(ratingmat, "realRatingMatrix")
```

#### Normalizing the data

```{r}
ratingmat_norm <- normalize(ratingmat)
```

#### Recommender Model using SVD (Singular Value Decomposition Method)

```{r}
recommender_model <- Recommender(ratingmat_norm, method = "SVD")

# Obtain top 10 recommendations for 5th user in data-set
recom <- predict(recommender_model, ratingmat[5], n=10) 

# Converting recommenderlab object to a readable list
recom_list <- as(recom, "list") 
```

#### Checking recommendations for the model

```{r}
recom_result <- matrix(0,10)

for (i in c(1:10)){
  recom_result[i] <- movies[as.integer(recom_list[[1]][i]),2]
}

# Printing top 10 recommendations
recom_result
```

#### Evaluating the model

```{r}
# Checking evaluation metrics for the SVD model using the 5-fold cross-validation
evaluation_scheme <- evaluationScheme(ratingmat, method="cross-validation", k=5, given=3, goodRating = 5) 

# Performing evaluation metrics for top 1,3,5, and 10 recommendations
evaluation_results <- evaluate(evaluation_scheme, method="SVD", n=c(1,3,5,10))

# Storing evaluation results 
eval_results <- getConfusionMatrix(evaluation_results)[[1]]

# Printing the evaluation results
eval_results
```

### Module 9: Case Study 2 - K-Means Algorithm

#### Loading required packages

```{r}
library('dplyr')
library('caTools')
```

#### Task-1 : Building the k-means clustering algorithm

```{r}
# a.Start off by extracting the ‘MonthlyCharges’, ‘tenure’ & ‘TotalCharges’ columns from the ‘customer_churn’ data.frame.Store the result in ‘customer_features’
# b.Remove any ‘NA’ values from ‘customer_features’ if present
# c.Build the k-means algorithm on top of ‘customer_features’.Here, the number of clusters should be 3
# d.Bind the clustering vector to ‘customer_features’.
# e.Extract observations belonging to individual clusters
```

#### Preparing the data

```{r}
# Loading the data-set
churn_4 <- read.csv('/home/carb0n/Desktop/r/d/customer_churn.csv')

# Checking structure of data-set
str(churn_4)

# Checking for NA values
colSums(is.na(churn_4))

# Removing "CustomerID" column as it is not needed
churn_4 <- within(churn_4, rm(customerID))

# Plotting box-plot for TotalCharges column to check for outliers
boxplot(churn_4$TotalCharges)

# Replacing NA values in TotalCharges column with mean value for TotalCharges column
churn_4$TotalCharges[is.na(churn_4$TotalCharges)] <- mean(churn_4$TotalCharges, na.rm = TRUE)

# Converting appropriate variables to factor-type 
as.factor(churn_4$gender) -> churn_4$gender
as.factor(churn_4$SeniorCitizen) -> churn_4$SeniorCitizen
as.factor(churn_4$Partner) -> churn_4$Partner
as.factor(churn_4$Dependents) -> churn_4$Dependents
as.factor(churn_4$PhoneService) -> churn_4$PhoneService
as.factor(churn_4$MultipleLines) -> churn_4$MultipleLines
as.factor(churn_4$InternetService) -> churn_4$InternetService
as.factor(churn_4$OnlineSecurity) -> churn_4$OnlineSecurity
as.factor(churn_4$OnlineBackup) -> churn_4$OnlineBackup
as.factor(churn_4$DeviceProtection) -> churn_4$DeviceProtection
as.factor(churn_4$TechSupport) -> churn_4$TechSupport
as.factor(churn_4$StreamingTV) -> churn_4$StreamingTV
as.factor(churn_4$StreamingMovies) -> churn_4$StreamingMovies
as.factor(churn_4$Contract) -> churn_4$Contract
as.factor(churn_4$PaperlessBilling) -> churn_4$PaperlessBilling
as.factor(churn_4$PaymentMethod) -> churn_4$PaymentMethod
as.factor(churn_4$Churn) -> churn_4$Churn
```

#### K-Means Model

```{r}
# Selecting specific columns to create data-frame
churn_4 %>% select("tenure", "MonthlyCharges", "TotalCharges") -> customer_features

# Building kmeans algorithm on top of data-frame
kmeans(customer_features$MonthlyCharges, 3) -> k_month

# Checking clustering vector values
head(k_month$cluster)

# Binding clustering vector values to column and converting to data-frame
cbind(Month = customer_features$MonthlyCharges, Cluster = k_month$cluster) -> month_group
as.data.frame(month_group) -> month_group
head(month_group)
```

#### Separating Clustering Vector Values

```{r}
# Separating clustering vector values for each cluster
month_group %>% filter(Cluster==1) -> month_group_1
month_group %>% filter(Cluster==2) -> month_group_2
month_group %>% filter(Cluster==3) -> month_group_3

# Checking separated cluster values for each cluster
head(month_group_1)
head(month_group_2)
head(month_group_3)
```