---
title: "R : Module 10 : Association Rule Mining"
output:
  html_document:
    df_print: paged
header-includes: \usepackage{fvextra} \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      tidy = 'formatR',
                      #tidy.opts = list(blank = FALSE, width.cutoff = 90),
                      prompt = F,
                      size = 'large')
```

## Module 10 : Case Study 1 - Apriori Algorithm

### Loading required packages

```{r}
library('recommenderlab')
library('arules')
library('arulesViz')
```

#### Viewing Groceries data-set

```{r}
data("Groceries")
```

#### Checking summary of data-set

```{r}
summary(Groceries)
```

#### First Apriori Algorithm

```{r}
# Building first apriori algorithm
rule1 <- apriori(Groceries, parameter = list(support=0.05, confidence=0.3))
```

#### Inspection Of First Apriori Algorithm Rules

```{r}
# Inspecting rule-1
inspect(head(rule1))

# Inspecting first five rules from rule1, sorting by lift
inspect(head(sort(rule1,by='lift'),5))
```

#### Plotting First Apriori Algorithm

```{r}
# Creating plot for rule-1
plot(rule1)

# Creating plot for rule-1 using method as 'grouped'
plot(rule1, method='grouped')
```

#### Second Apriori Algorithm

```{r}
# Building second apriori algorithm
rule2 <- apriori(Groceries, parameter = list(support=0.004,confidence=0.7))
```

#### Inspection Of Second Apriori Algorithm Rules

```{r}
# Inspecting rule2
inspect(head(rule2))

# Inspecting first five rules from rule1, sorting by lift
inspect(head(sort(rule2,by='lift'),5))
```

#### Plotting Second Apriori Algorithm

```{r}
# Creating plot for rule2
plot(rule2)

# Creating plot for rule2 using method as 'grouped'
plot(rule2, method='grouped')
```

## Module 10 : Case Study 2 - Recommender System

### Loading required packages

```{r}
library('dplyr')
library('caTools')
library('recommenderlab')
```

#### Loading MovieLense data-set from recommenderlab package

```{r}
data(MovieLense)
```

#### Checking structure of data-set 

```{r}
str(MovieLense)
```

#### Viewing Movie-Names from data-set

```{r}
head(MovieLense@data@Dimnames[[2]])
```

### Preparing the data

#### Removing records with 0 ratings

```{r}
# Storing ratings data as a vector 
vector_ratings <- as.vector(MovieLense@data)

# Checking unique values of ratings
unique(vector_ratings)

# Removing records having 0 ratings
vector_ratings <- vector_ratings[vector_ratings!=0] 
```

#### Checking distribution of ratings

```{r}
table_ratings <- table(vector_ratings)
table_ratings
```

#### Filtering data from original MovieLense data-set

```{r}
sample_movie <- MovieLense[rowCounts(MovieLense) >= 100, colCounts(MovieLense) >= 120]
```

#### Splitting the data-set

```{r}
# Creation of sample from filtered data
set.seed(1)
split_movie <- sample(x=c(T,F), size=nrow(sample_movie), replace=T, prob=c(0.8,0.3))

# Creation of train and test sets
recc_train <- sample_movie[split_movie,] 
recc_test <- sample_movie[!split_movie,]
```

### (A) User based Collaborative Filtering (UBCF)

#### Model Creation

```{r}
# Building Model
ubcf_model <- Recommender(data=recc_train, method='UBCF')
```

#### Recommending on test-data

```{r}
# Number of Recommendations to be made
n_recommended_ubcf <- 10

# Prediction for test data
recc_predicted_ubcf <- predict(object=ubcf_model, newdata=recc_test, n=n_recommended_ubcf)
```

#### Viewing Recommendations for UBCF Model

```{r}
# Recommendations for User 3
user_3_movie_numbers <- recc_predicted_ubcf@items[[3]]
recc_predicted_ubcf@itemLabels[user_3_movie_numbers]

# Recommendations for User 5
user_5_movie_numbers <- recc_predicted_ubcf@items[[5]]
recc_predicted_ubcf@itemLabels[user_5_movie_numbers]
```

### (B) Item-Based Collaborative Filtering(IBCF)

#### Model Creation

```{r}
# Building Model
ibcf_model <- Recommender(data=recc_train, method='IBCF')  
```

#### Recommending on test-data

```{r}
# Number of Recommendations to be made
n_recommended_ibcf <- 12

# Prediction for test data
recc_predicted_ibcf <- predict(object=ibcf_model, newdata=recc_test, n=n_recommended_ibcf)
```

#### Viewing Recommendations for IBCF Model

```{r}
# Recommendations for User 3
user_3_movie_numbers <- recc_predicted_ibcf@items[[3]]
recc_predicted_ibcf@itemLabels[user_3_movie_numbers]

# Recommendations for User 5
user_5_movie_numbers <- recc_predicted_ibcf@items[[5]]
recc_predicted_ibcf@itemLabels[user_5_movie_numbers]
```